---
title: "will_Lab10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Simple rules to judge a question

We are curious about what make the question or answer in the dataset to be a good/bad one. 

Straightforwardly, we find two  coarse features, *i.e.* the length of the title and body of the questions. 

```{r, echo = TRUE}
stackQuestions <- read.csv("D:\\Desktop\\Questions_trunc.csv",header= TRUE, sep = ",")
stackScore_Questions <- stackQuestions$Score
countBody_Questions <- sapply(gregexpr("\\W+", stackQuestions$Body), length)
countTitle_Questions <- sapply(gregexpr("\\W+", stackQuestions$Title), length)

df <- data.frame(countTitle_Questions, stackScore_Questions)
df2 <- data.frame(countBody_Questions, stackScore_Questions)


ggplt <- ggplot(df, aes(x = countTitle_Questions, y = stackScore_Questions)) + geom_point()

ggplt2 <- ggplot(df, aes(x = countBody_Questions, y = stackScore_Questions)) + geom_point()



ggplt
```
It coincides with our first thought, a good question is with a balanced length of title.
```{r}
ggplt2
```

On the other side, from the graph above, we can draw the conclusion that the quanlity of a question has a negative correction with the length of body of the question.

But the feature is too coarse to character how to be a good question.

## Simple rules to judge an answer
Intuitively, the answer is too long or, too short, it may not be good enough. 

```{r, echo = TRUE}

stackAnss <- read.csv("D:\\Desktop\\Answers_trunc.csv", header = TRUE, sep = ",")
score_anss <- stackAnss$Score
countBody_Ans <- sapply(gregexpr("\\W+", stackAnss$Body), length)
df3 <- data.frame(countBody_Ans, score_anss)

ggplt3 <- ggplot(df3, aes(x = countBody_Ans, y = score_anss)) + geom_point()
ggplt3
```

From the plot, we can see most of the answers are less than 1000 words. From practical point of view, too long meeans impatience. So a balanced length of the answer can improve the quality of the answer. Elegancy and simpleness can earn a good score.

But again this feature is too coarse to character how to be a good answer.