---
title: "Assignment 11"
author: "David Du"
date: "4/5/2018"
output: html_document
---

```{r setup, include=FALSE}
library(babynames)
library(stringr)
library(tidyverse)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE)
```

First three letters of your names

sum(str_count(babynames$names),"dav")   # 274
sum(str_count(babynames$names),"Dav")   # 3852
3852/(274+3852)= 0.934

Ratio of "Dav" = 0.934

How many versions of the female name “Ariel” were there in 1973? 1988? 1990?

```{r , echo=FALSE}
rachel1973<-filter(babynames, str_detect(babynames$name, "Ra[e]?ch.*l"),year==1973, sex=="F") %>% count()
rachel1988<-filter(babynames, str_detect(babynames$name, "Ra[e]?ch.*l"),year==1988, sex=="F") %>% count()
rachel1990<-filter(babynames, str_detect(babynames$name, "Ra[e]?ch.*l"),year==1990, sex=="F") %>% count()
```

How many versions for the female name “Rachel” for those years?

```{r, echo=FALSE}
ariel1973<-filter(babynames, str_detect(babynames$name, "Ar[iy]+.l[^a]?$"),year==1973, sex=="F") %>% count()
ariel1988<-filter(babynames, str_detect(babynames$name, "Ar[iy]+.l[^a]?$"),year==1988, sex=="F") %>% count()
ariel1990<-filter(babynames, str_detect(babynames$name, "Ar[iy]+.l[^a]?$"),year==1990, sex=="F") %>% count()
```

What are the chances a girl born in 1973 would be named either Rachel or Ariel (including various versions)? In 1988? In 1990?

```{r pressure, echo=FALSE}
Girls1973<- filter(babynames,year==1973, sex=="F") %>% count()
Chance1973<-(rachel1973+ariel1973)/Girls1973
Chance1973

Girls1988<- filter(babynames,year==1988, sex=="F") %>% count()
Chance1988<-(rachel1988+ariel1988)/Girls1988
Chance1988

Girls1990<- filter(babynames,year==1973, sex=="F") %>% count()
Chance1990<-(rachel1990+ariel1990)/Girls1990
Chance1990

```

“Did The Little Mermaid cause more baby girls to be named Ariel (and various spellings)?
```{r}
filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year==1988,sex=="F") %>% count(wt=prop) #0.000841196	

filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year==1990,sex=="F") %>% count(wt=prop) #0.002612356	
```


What was the change in proportion in female Ariel baby names?

Here we can use the ratio from 1988 to 1990 to calculate the change as following 0.000841196-0.002612356= -0.00177116. Therefore, we could conclude that the Little Mermaid cause less baby girls to be named Ariel 


Considering all the changes in female names starting with a vowel from 1988 to 1990?
```{r}

Agirls1 <- filter(babynames,str_detect(babynames$name,"A"),year==1988,sex=="F") %>% count() # 1238
Agirls2 <- filter(babynames,str_detect(babynames$name,"A"),year==1990,sex=="F") %>% count() # 1458
Egirls1 <- filter(babynames,str_detect(babynames$name,"E"),year==1988,sex=="F") %>% count() # 386
Egirls2 <- filter(babynames,str_detect(babynames$name,"E"),year==1990,sex=="F") %>% count() # 422
Ogirls1 <- filter(babynames,str_detect(babynames$name,"O"),year==1988,sex=="F") %>% count() # 66
Ogirls2 <- filter(babynames,str_detect(babynames$name,"O"),year==1990,sex=="F") %>% count() # 74
Igirls1 <- filter(babynames,str_detect(babynames$name,"I"),year==1988,sex=="F") %>% count() # 132
Igirls2 <- filter(babynames,str_detect(babynames$name,"I"),year==1990,sex=="F") %>% count() # 175
Ugirls1 <- filter(babynames,str_detect(babynames$name,"U"),year==1988,sex=="F") %>% count() # 12
Ugirls2 <- filter(babynames,str_detect(babynames$name,"U"),year==1990,sex=="F") %>% count() # 17

```
 
What percentile is the change in Ariel names?
```{r}
Ariel.diff <- filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year==1990,sex=="F") %>% count(wt=prop) - filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year==1988,sex=="F") %>% count(wt=prop)

filter(Agirls,diff<Ariel.diff) %>% arrange(desc(diff))
 filter(Agirls,diff<Ariel.diff) %>% count() /
  Agirls %>% count()  #100 percentile

```


In this case, the variable of interest is the change in proportion of female “Ariel” baby names from 1988 to 1990. It’s hard to make conclusions if you only know how one variable changes over time. What about the changes of all the other names? Is the change in the proportion of Ariels big, medium, or small compared to the changes in other female names over that time period? Create a plot that shows the change in Ariel female names from 1988 to 1990 in the context of changes in proportions for other female names over that same time period. How does the change in Ariel names compare? Describe your plot in words and include answers to the questions in this paragraph in your description of your plot.

```{r}
Agirls <- inner_join(Agirls1,Agirls2,by="name") %>% mutate(diff=prop.y-prop.x) %>% select(name,prop.x,prop.y,diff)
ggplot(Agirls) +
  geom_histogram(aes(x=diff),bins=100) + 
  geom_vline(aes(xintercept=Ariel.diff),color="red",lwd=1) 
```

We could find that the change of the ariel names is pretty big since for most of the names, the change has been around 0 but for the histogram I have drawn for Ariel. It is the far mostright so it rrepresents that it has biggest change.


How did the popularity of your name (in the US) change from the year you were born to 2015? How did the popularity of your name change from X years before your birth year to your birth year (X is 2015 minus your birth year)? Use a regex to combine at least two versions of your name.


```{r}
David1<- filter(babynames,str_detect(babynames$name,"^David"),year==1977,sex=="M") %>% count(wt=prop) 
David2<- filter(babynames,str_detect(babynames$name,"^Dav"),year==1977,sex=="M") %>% count(wt=prop) 
David3<- filter(babynames,str_detect(babynames$name,"^David"),year==1996,sex=="M") %>% count(wt=prop) 
David4<- filter(babynames,str_detect(babynames$name,"^Dav"),year==1996,sex=="M") %>% count(wt=prop) 
David5<- filter(babynames,str_detect(babynames$name,"^David"),year==2015,sex=="M") %>% count(wt=prop) 
David6<- filter(babynames,str_detect(babynames$name,"^Dav"),year==2015,sex=="M") %>% count(wt=prop) 

David1977.diff<- filter(babynames,str_detect(babynames$name,"^David"),year==1977,sex=="M") %>% count(wt=prop) - filter(babynames,str_detect(babynames$name,"^Dav"),year==1977,sex=="M") %>% count(wt=prop)  # -0.000442135
David1996.diff<- filter(babynames,str_detect(babynames$name,"^David"),year==1996,sex=="M") %>% count(wt=prop)- filter(babynames,str_detect(babynames$name,"^Dav"),year==1996,sex=="M") %>% count(wt=prop)  # -0.001253551
David2015.diff<- filter(babynames,str_detect(babynames$name,"^David"),year==2015,sex=="M") %>% count(wt=prop) - filter(babynames,str_detect(babynames$name,"^Dav"),year==2015,sex=="M") %>% count(wt=prop)  # -0.001097257

```


Two histograms to show the changes. 
```{r}
ALL1977<-filter(babynames,year==1977)
ALL1996<-filter(babynames,year==1996)
ALL2015<-filter(babynames,year==2015)
ALL96TO15 <- inner_join(ALL1996,ALL2015,by="name") %>% mutate(diff=prop.y-prop.x) %>% select(name,prop.x,prop.y,diff)
ggplot(ALL96TO15) +  geom_histogram(aes(x=diff),bins=100) + geom_vline(aes(xintercept=David2015.diff),color="red",lwd=1)

```

```{r}
ALL77TO96 <- inner_join(ALL1977,ALL1996,by="name") %>% mutate(diff=prop.y-prop.x) %>% select(name,prop.x,prop.y,diff)
ggplot(ALL77TO96) +  geom_histogram(aes(x=diff),bins=100) + geom_vline(aes(xintercept=David1977.diff),color="red",lwd=1)
```

The 1st histogram shows the change of the proportion of my name between 1997 and 2015. During this time I have found that my name has been decreased by 0.019% which is below the change compared with other baby names. The 2nd histogram shows the change of the proportion of my name between 1979 and 1997. During this time I have found that my name has been increased by 0.087% which is above the change compared with other baby names.  
