---
title: "AbbyLab11"
author: "Abby"
date: "April 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(babynames)
```
## First Three Letters of Names

```{r why}
Abb<-filter(babynames,str_detect(babynames$name,"^Abb"))
abb<-filter(babynames, str_detect(babynames$name,"abb"))
abbtotal<-sum(abb$n)
Abbtotal<-sum(Abb$n)
Abbprop<-Abbtotal/(abbtotal+Abbtotal)
```
Overall, the ratio of names in the dataset that start with the first three letters of Abby divided by total names that contain (in order) the first three letters of the Abby is 0.96522, or about 96.52%, which means most names with abb in them also begin with Abb.

```{r over time}
Abb88<-filter(babynames,str_detect(babynames$name,"^Abb"),year==1988,sex=="F") %>% count(wt=prop) 
```

## Ariel and Rachel Regexs
```{r need it}
Ariel1<-filter(babynames, str_detect(babynames$name, "Ar[iy]+.l+[^a]?$"),sex=="F")
A73<-filter(Ariel1, year==1973, sex=="F")
##2 versions of Ariel in 1973
A88<-filter(Ariel1, year==1988,sex=="F")
##7 versions of Ariel in 1988
A90<-filter(Ariel1, year==1990, sex=="F")
##8 versions of Ariel in 1990

Rachel1<-filter(babynames, str_detect(babynames$name, "Ra[ey]?ch.*l"),sex=="F")
R73<-filter(Rachel1,year==1973)
##13 versions of Rachel
R88<-filter(Rachel1, year==1988)
##19 versions of Rachel
R90<-filter(Rachel1,year==1990)
##17 versions of Rachel

n73<-filter(babynames,year==1973)
chance73<-(sum(R73$n)+sum(A73$n))/sum(n73$n)
## Chances of being named Rachel or Ariel in 1973 0.00278237

n88<-filter(babynames,year==1988)
chance88<-(sum(R88$n)+sum(A88$n))/sum(n88$n)
## Chances of being named Rachel or Ariel in 1988 0.00567

n90<-filter(babynames, year==1990)
chance90<-(sum(R90$n)+sum(A90$n))/sum(n90$n)
## Chances of being named Rachel or Ariel in 1990 0.00612
```

## The Little Mermaid Effect
```{r littlemermaid}
## Finding the difference for Ariel
filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year==1988,sex=="F") %>% count(wt=prop)# 0.000841196
filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year==1990,sex=="F") %>% count(wt=prop) #0.002612356
vowelgirls1 <- filter(babynames,str_detect(babynames$name,"^[AEIOU]"),year==1988,sex=="F")
vowelgirls2 <- filter(babynames,str_detect(babynames$name,"^[AEIOU]"),year==1990,sex=="F")
vowelgirls <- inner_join(vowelgirls1,vowelgirls2,by="name") %>% mutate(diff=prop.y-prop.x) %>% select(name,prop.x,prop.y,diff)
Ariel.diff <- filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year==1990,sex=="F") %>% count(wt=prop) - filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year==1988,sex=="F") %>% count(wt=prop)
ggplot(vowelgirls) +
  geom_histogram(aes(x=diff),bins=100) + 
  geom_vline(aes(xintercept=Ariel.diff),color="blue",lwd=1) 
```
From the plot above we can see that the change in Ariel names over between 1988 and 1990 is much bigger than the changes in other girls names beginning with vowels over the same time period. Thus, from this it seems as the the movie The Little Mermaid did lead to an increase in the proportion of girls named some version of Ariel between the years 1988 and 1990.

## Abby name
```{r Abbyprops}
## Dr. Vance way
filter(babynames,str_detect(babynames$name,"^Abb[iy]$"),year==1997,sex=="F") %>% count(wt=prop)# 0.0008440914
filter(babynames,str_detect(babynames$name,"^Abb[iy]$"),year==2015,sex=="F") %>% count(wt=prop) #0.0003839577
filter(babynames,str_detect(babynames$name,"^Abb[iy]$"),year==1979,sex=="F") %>% count(wt=prop) #0.0004787982
all97 <- filter(babynames,year==1997)
all2015 <- filter(babynames,year==2015)

Abby.diff15 <- filter(babynames,str_detect(babynames$name,"^Abb[iy]$"),year==2015,sex=="F") %>% count(wt=prop) - filter(babynames,str_detect(babynames$name,"^Abb[iy]$"),year==1997,sex=="F") %>% count(wt=prop)

all97to15 <- inner_join(all97,all2015,by="name") %>% mutate(diff=prop.y-prop.x) %>% select(name,prop.x,prop.y,diff)
ggplot(all97to15) +
  geom_histogram(aes(x=diff),bins=100) + 
  geom_vline(aes(xintercept=Abby.diff15),color="red",lwd=1)
```
Above is the change in proportion of my name between 1997 and 2015. Between 1997 and 2015 the popularity of my name has decreased by 0.046% which is below the change in proportion of most other baby names.

```{r Abbyprops from 1979}
all79 <- filter(babynames,year==1979)

Abby.diff79 <- filter(babynames,str_detect(babynames$name,"^Abb[iy]$"),year==1997,sex=="F") %>% count(wt=prop) - filter(babynames,str_detect(babynames$name,"^Abb[iy]$"),year==1979,sex=="F") %>% count(wt=prop)
Abby.diff792 <- filter(babynames,str_detect(babynames$name,"^Abby$"),year==1997,sex=="F") %>% count(wt=prop) - filter(babynames,str_detect(babynames$name,"^Abby$"),year==1979,sex=="F") %>% count(wt=prop)

all79to97 <- inner_join(all79,all97,by="name") %>% mutate(diff=prop.y-prop.x) %>% select(name,prop.x,prop.y,diff)
ggplot(all79to97) +
  geom_histogram(aes(x=diff),bins=100) + 
  geom_vline(aes(xintercept=Abby.diff79),color="blue",lwd=1)
```
Above is the change in proportion of my name between 1979 and 1997.Between 1979 and 1997 the popularity of my name has increased by 0.037% which is above the change in proportion of most other baby names.