---
title: "Lab 4"
author: "WADDS"
date: "February 7, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz='bwt/oz')
```

## Guiding Question
We want to know if smoking during pregancy affects the gestation period and the birth weights of the baby. We want to know what confounding variables may be influencing the relationship between gestation period and smoking and birth weight and smoking, to determine if there is a relationship between these variables.

## Recommendations

## Preliminary Question
Which variables are similar among smokers and non-smokers? This is what we will be analyzing in our individual parts below.

## Mother's Height to Weight Ratio

In order to analyze possible factors that might contribute to the likelihood of a baby being premature, I analyzed the mother's height to weight ratio, separating the results into mothers who had premature babies (right) and mothers who did not (left). The plot of these results are shown below. 


```{r plot, echo=FALSE}
babies$Premature<-factor(babies$Premature)
babies<-subset(babies, Premature==0 | Premature==1)
ggplot(data=babies,aes(x=mht,y=mpregwt))+ggtitle("Mother's Height to Weight Ratio")+geom_jitter()+geom_smooth(se=FALSE,method="lm")+facet_wrap(~Premature)+geom_abline(slope=5, intercept=-200,color="red")
```


The red line indicates a generally accepted standard of what a healthy woman would weigh close to for her given height. This baseline was created from the estimate on livestrong.com. It is improtant to note that this baseline is general, and what is considered healthy varies from person to person. The blue line is the linear approximation the mothers' height to weight ratio within the data. 

Though both mothers with premature babies and mothers with non-premature babies have a variety of height to weight ratios, including some that seem to vary significantly from the "healthy" line, there is no significant different in the trends of the mothers with premature babies and the mother of this. Because of this, I would conclude that the mother's height to weight ratio does not have much of an affect on whether or not the child is premature. 

## Father's Weight to Birthweight

Since most children take after their parents in height once they are fully grown, it would be interesting to note if birthweight is influenced by the weight of the father. If bigger babies have heavier fathers, then this may be a confounding variable in regards to birthweight and smoking. The question I hope to answer is, do father's weight and birthweight have a strong correlation?

```{r abby}
ggplot(data=babies, mapping=aes(x=dwt, y=bwtoz))+geom_jitter()+geom_smooth(re=FALSE, method="lm")
```
This plot illustrates how dad's weight influences a babie's birthweight. From the plot there does not seem to be much of a relationship between a dad's weight and the birthweight of the baby. The relationship would most likely be moderate positive, as there does seem to generally be a trend that as a dad's weight increases, so does the babie's. This could be because birthweight is more influenced by a mother's weight before pregnancy, but unfortunately we do not have data to test this observation. One interesting factor we could test, is if there appears to be a stronger relationship between Dad's weight and birthweight when we account for if the mother smoked during pregnancy.
```{r abby}
ggplot(data=babies, mapping=aes(x=dwt, y=bwtoz))+geom_jitter()+facet_wrap(~ifelse(smoke,"Doesn't Smoke", "Smokes"))+geom_smooth(re=FALSE, method="lm")
```
From what we see in the plots, among those who don't smoke, there appears to be a strong, positive relationship between the Dad's weight and the birthweight, while there seems to be a weak correlation between a Dad's weight and a babie's weight among those who do smoke. This could indicate that a mother's smoking habits may be a confounding variable impacting the influence of the Dad's weight on the birthweight. This could indicate that smoking decreases the birthweight, regardless of the Father's weight beforehand.


## Individual Contributions
Abby: I created a plot to determine if a babie's birthweight was impacted by the Father's weight. From those results I then created another set of plots to see if smoking made the relationship between Dad's weight and birthweight stronger or weaker. 
