---
title: "Lab 4"
author: "WADDS"
date: "February 7, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz='bwt/oz')
```

## Guiding Question



```{r cars}
summary(cars)
```

## Plots and Analysis

Since most children take after their parents in height once they are fully grown, it would be interesting to note if birthweight is influenced by the weight of the father. If bigger babies have heavier fathers, then this may be a confounding variable in regards to birthweight and smoking. The question I hope to answer is, do father's weight and birthweight have a strong correlation?

```{r abby}
ggplot(data=babies, mapping=aes(x=dwt, y=bwtoz))+geom_jitter()+geom_smooth(re=FALSE, method="lm")
```
This plot illustrates how dad's weight influences a babie's birthweight. From the plot there does not seem to be much of a relationship between a dad's weight and the birthweight of the baby. The relationship would most likely be moderate positive, as there does seem to generally be a trend that as a dad's weight increases, so does the babie's. This could be because birthweight is more influenced by a mother's weight before pregnancy, but unfortunately we do not have data to test this observation. One interesting factor we could test, is if there appears to be a stronger relationship between Dad's weight and birthweight when we account for if the mother smoked during pregnancy.
```{r abby}
ggplot(data=babies, mapping=aes(x=dwt, y=bwtoz))+geom_jitter()+facet_wrap(~ifelse(smoke,"Doesn't Smoke", "Smokes"))+geom_smooth(re=FALSE, method="lm")
```
From what we see in the plots, among those who don't smoke, there appears to be a strong, positive relationship between the Dad's weight and the birthweight, while there seems to be a weak correlation between a Dad's weight and a babie's weight among those who do smoke. This could indicate that a mother's smoking habits may be a confounding variable impacting the influence of the Dad's weight on the birthweight. This could indicate that smoking decreases the birthweight, regardless of the Father's weight beforehand.
## Recommendations

## Preliminary Question

## Individual Contributions
Abby: I created a plot to determine if a babie's birthweight was impacted by the Father's weight. From those results I then created another set of plots to see if smoking made the relationship between Dad's weight and birthweight stronger or weaker. 
