---
title: "Assignment 9"
author: "David Du"
date: "3/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(nycflights13)
library(forcats)
library(lubridate)
library(stringr)

```

## Chapter 13 

# Problem: 
What does it mean for a flight to have a missing tailnum? What do the tail numbers that don’t have a matching record in planes have in common? (Hint: one variable explains ~90% of the problems.)

# Solution:

```{r }
anti_join(flights, planes, by = "tailnum") %>% 
count(carrier, sort=TRUE)
```


Here, I have anti_join function which could find out the unmatched record, here we need to find out flights have missing tailnum. Therefore, I use this function to find out Amercian Airline and Envoy Airline have most missing record for tailnum. Missing tailnum could mean that the flight is unregistered, unlicensed flight.


## Chapter 14

# Problem and Solution
Create regular expressions to find all words that:

Start with a vowel.

```{r, echo=FALSE}
#str_view(stringr::words, "^[aeiou]", match = TRUE)
```

That only contain consonants. (Hint: thinking about matching “not”-vowels.)

```{r, echo=FALSE}
#str_view(stringr::words, "^[^aeiou]+$", match=TRUE)
```

End with ed, but not with eed.

```{r, echo=FALSE}
str_view(stringr::words, "^ed$|[^e]ed$", match = TRUE)
```

End with ing or ise.

```{r, echo=FALSE}
#str_view(stringr::words, "i(ng|se)$", match = TRUE)
```

For some of problems, the output is quite long, so I just leave my code.

## Chapter 15

# Problem 
There are some suspiciously high numbers in tvhours. Is the mean a good summary?
```{r , echo=FALSE}
summary(gss_cat[["tvhours"]])
ggplot(gss_cat, aes(x=tvhours),filter(!is.na(tvhours)))+geom_histogram(binwidth=1)
```

I think use median in this question could be much more convince than mean, since there are some values are extremely large, and we could see that there are a lot of people watching TV for 2.5 hours. However, the hours of TV does not look like surprising me. 

## Chapter 16

Functions need to be used prior to do exercises

```{r , echo=FALSE}
make_datetime_100 <- function(year, month, day, time) {
make_datetime(year, month, day, time %/% 100, time %% 100)
}
flights_dt <- flights %>%
filter(!is.na(dep_time), !is.na(arr_time)) %>%
mutate(
dep_time = make_datetime_100(year, month, day, dep_time),
arr_time = make_datetime_100(year, month, day, arr_time),
sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)
) %>%
select(origin, dest, ends_with("delay"), ends_with("time"))

```


# Problem 
What makes the distribution of diamonds$carat and flights$sched_dep_time similar?

# Solution
```{r, echo=FALSE}
ggplot(diamonds, aes(x = carat)) + geom_density()
ggplot(diamonds, aes(x = carat %% 1 * 100)) +geom_histogram(binwidth = 1)
ggplot(flights_dt, aes(x = minute(sched_dep_time))) +geom_histogram(binwidth = 1)
```

Here, I have draw three graph in order to help us understand the relationship between these variables. First, we could take out the flights time between 0 and 30 minutes. We can clearly see that the distribution of the flights depart is similar with the count of carat and diamonds. Here, I have divided the flights time by 5 minutes, so we could draw graph based on ending with 5, 0 minutes. This corresponds the carats with 0, 1/3,1/2, 2/3. I have multiply with 100 to see visualization. These trends and numbers could be made by human factors and it is also makes much sense. 


## Summary of Contribution

David: For chapter 13, I have learnt anti_join function in order to find out the number of flight that has missing record of tailnum. For chapter 14, I have learnt the ways to filter with strings with specific requirements. For chapter 15, I have draw historgam in order to visualize the TV hours. In chapter 16, I have manuipulated with time and dates to answer the question. In this lab, I have implemented with ggplot, geom_density, and geom_histogram function.




