---
title: "Will_Lab9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13) #for Ex.13.4.6
library(stringr) # for Ex.14.2.5#2
library(forcats) #included for Ex.15.4.1
library(lubridate) # included for the last excercise
```


## EX. 13.4.6 #3
> *Is there a relationship between the age of a plane and its delays?*

### Solution 1
Counter to our intuition, The minimal time of departure dealay tends to increase slightly with the aging of the plane.
```{r, echo = TRUE}
### Specify the the ages of the planes.
ages_plane <- 
  planes %>%
  mutate(age = 2018 - year) %>%
  select(tailnum, age)

### Minimal minutes of delay
flights %>%
  inner_join(ages_plane, by = "tailnum") %>%
  group_by(age) %>%
  ### Here we use the average 
  filter(!is.na(dep_delay)) %>%
  summarise(delay_min = min(dep_delay)) %>%
  ggplot(aes(x = age, y = delay_min)) +
      geom_point() +
      geom_line()

### Maximal minutes of delay
flights %>%
  inner_join(ages_plane, by = "tailnum") %>%
  group_by(age) %>%
  ### Here we use the average 
  filter(!is.na(dep_delay)) %>%
  summarise(delay_max = max(dep_delay)) %>%
  ggplot(aes(x = age, y = delay_max)) +
      geom_point() +
      geom_line()
```

But suprisingly, the maximal minutes of departure delays decrease with the aging of the plane, especially when the
the plane is older than 20 years.

## Ex. 13.3.3.1 #1
> Create regular expressions to find all words that:

> Start with a vowel.
That only contain consonants. (Hint: thinking about matching "not"-vowels.)
End with `ed`, but not with `eed`.
End with `ing` or `ise`.

### Solution 2
```{r}
str_c("a", c("a","b"), sep = " ", collapse = ",")
str_c("a", c("a","b"), sep = " ")
```


## Ex. 15.4.1 #2
> *For each factor in `gss_cat` identify whether the order of the levels is arbitrary or principled.*

### Solution 3
There are five six categorical variables: `marital`, `race`, `rincome`, `partyid`, `relig`, `denom`.

The ordering of martial is *partially* principled.  It can be grouped by "never married", married in the past (separated, divorced, widowed), and "married"

```{r, echo = TRUE}
keep(gss_cat, is.factor) %>% names()
levels(gss_cat[["marital"]])

gss_cat %>%
  ggplot(aes(x = marital)) +
  geom_bar()

### Obviously, The ordering of race is principled.

levels(gss_cat$race)

gss_cat %>%
  ggplot(aes(race)) +
  geom_bar(drop = FALSE)


###  `rincome` is principledly ordered if we don't take "No answer", "Don't know", "Refused" and "Not applicable" into account, otherwise, it is arbitary.
levels(gss_cat$rincome)

### The levels of `relig` is arbitrary.
gss_cat %>%
  ggplot(aes(relig)) +
  geom_bar() +
  coord_flip()
```


## Ex. 16.3.4 #4

> *How does the average delay time change over the course of a day? Should you use `dep_time` or `sched_dep_time`? Why?*

### Solution 4
`sched_dep_time` is relevant, because the delay time varies with it. So it is used. But `dep_time` is not used explictly here, it is only usded to clean out the missing data.
```{r, echo = TRUE}
### Convert the time over the day
make_datetime1 <- function(year, month, day, time) {
  make_datetime(year, month, day, time %/% 100, time %% 100)
}

flights1 <- flights %>% 
  filter(!is.na(dep_time), !is.na(arr_time)) %>% 
  mutate(
    dep_time = make_datetime1(year, month, day, dep_time),
    arr_time = make_datetime1(year, month, day, arr_time),
    sched_dep_time = make_datetime1(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime1(year, month, day, sched_arr_time)
  ) %>% 
  select(origin, dest, ends_with("delay"), ends_with("time"))


flights1 %>%
  mutate(sched_dep_hour = hour(sched_dep_time)) %>%
  group_by(sched_dep_hour) %>%
  summarise(dep_delay = mean(dep_delay)) %>%
  ggplot(aes(y = dep_delay, x = sched_dep_hour)) +
  geom_point() +
  geom_smooth()
```
