---
title: "Will_lab 14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2) 
library(dplyr) 
library(forcats) 
#install.packages("corrplot")
library(corrplot) 
#install.packages("GGally")
library(GGally) 
#install.packages("ggridges")
library(ggridges)
library(modelr)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)
```

## Prediction of Ages

We use the weight as a independent variable, to predict the age.

```{r, echo = TRUE}
mod_e1 <- lm(age ~ weight, data = cdc)
mod_e2 <- lm(age ~ poly(weight, 2), data = cdc)
mod_e3 <- lm(age ~ poly(weight, 3), data = cdc)
cdc %>% 
    expand(weight) %>% 
    gather_predictions(mod_e1, mod_e2, mod_e3) %>% 
    ggplot(aes(weight, pred, colour = model)) +
    geom_point() + 
    geom_line()
```

From the plot above, we see that both the models of polynomials of degree 2 and 3 fit the trend well, in comparison with the the plots previously investigated in Lab 3.

## Significance

As we learn more statistical techniques, we can study the relationship between weight and smoke more closely. For example, we can figure out whether the positive association between weight and `smoke100` is significant.


We take a naive approach, just compare the mean values of `smoke100` with weight ranging between 0~100 and 100~200 

### Null Hypothesis
Assume that the averages of `smoke100` are equal.

```{r, echo=TRUE}
# Weight in 0~100
smoke_weightless100 <- filter(cdc,weight < 100 & gender == "m") %>% select(smoke100) %>% unlist
average_smoke_weightless100 <- mean(smoke_weightless100)

# Weight in 100~200
smoke_weightless200 <- filter(cdc,weight >= 100 & weight < 200 & gender == "m") %>% select(smoke100) %>% unlist
average_smoke_weightless200 <- mean(smoke_weightless200)

# Observed difference
obsDiff <- average_smoke_weightless200 - average_smoke_weightless100
obsDiff

```

### Mix Up
Now assume that the weight labels don't matter. We mix up the labels, to calculate the test statistic for the mixed-up labels.


```{r, echo = TRUE}
N = 12
diffAverage <- replicate(1000, {
    all <- sample(c(smoke_weightless100,smoke_weightless200))
    testGroup1 <- all[1:N]
    testGroup2 <- all[(N+1):(2*N)]
  return(mean(testGroup2) - mean(testGroup1))
})
hist(diffAverage)
abline(v=obsDiff, col="red", lwd=2)
```

### The Percentile
Now we compare the real data summary statistic with the distribution generated above. 
```{r, echo = FALSE}
# minDiff <- min(diffAverage)
# maxDiff <- max(diffAverage)
# prob <- (obsDiff - minDiff) / (maxDiff - minDiff)
```
```{r, echo = TRUE}
# Find the percentile
percentile <- ecdf(diffAverage)
percentile(obsDiff)
```

From the distribution, the observed value, *i.e.*, the difference of the two means are not "*special*" among the 
sample when mixing up the weight labels. we have to accept the null hypothesis. 

In a word, the relationship between weight and smoke for the people with healthy plan is not statistically significant as we previously thought.

One more thing, from the work above, we learn that, sometimes the coarse graphic analysis combining educated guess is misguiding. With more sophisticated tool used, we can understand the numbers better and statistics will find more difficult to lie to us.