---
title: "Lab 4"
author: "Xinzhi Rao"
date: "2/9/2018"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) #renaming `bwt/oz` because the "/" sign requires special care
```


```{r echo=FALSE}
ggplot(data=babies,aes(x=gestation,y=bwtoz))+geom_smooth(se=FALSE,method=lm)+geom_point()+facet_wrap(~smoke)+scale_y_continuous(limits=c(0,190))
```

## Does smooke affect the relationship between gestation and bwtoz?
####  By looking the plot, we will find that when people do not smoke, they will have a normal slope relationship between gestation and bwtoz. Most of bwt are between 112.5 to 137.5, which is a normal range. But for smoking people, the relationship between gestation and bwtoz will have a sharp slope, it means, when the getation longer, bwtoz will increase a lot. Sometimes bwt will exceed the normal range. This definitely shows that smoking affects the relationship between gestation and bwtoz. It also leads to negative relationship.

## Contribution
#### I created a plot to see the whether smoking could affect the relatioinship between gestation and bwtoz and also to see whether the smoking leads to positive or negative impact.



