---
title: "Lab 13 WADDS"
author: "WADDS"
date: "April 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
sugardata<-read.csv("sugar_consumption2.csv")
sugardata%>%
  setnames(old=c("X1961","X1962","X1963","X1964","X1965","X1966","X1967","X1968","X1969","X1970","X1971","X1972","X1973","X1974","X1975","X1976","X1977","X1978","X1979","X1980","X1981","X1982","X1983","X1984","X1985","X1986","X1987","X1988","X1989","X1990","X1991","X1992","X1993","X1994","X1995","X1996","X1997","X1998","X1999","X2000","X2001","X2002","X2003","X2004"), new=c("1961","1962","1963","1964","1965","1966","1967","1968","1969","1970","1971","1972","1973","1974","1975","1976","1977","1978","1979","1980","1981","1982","1983","1984","1985","1986","1987","1988","1989","1990","1991","1992","1993","1994","1995","1996","1997","1998","1999","2000","2001","2002","2003","2004"))      
sugardata3<-sugardata%>%
  gather('1961','1962','1963','1964','1965','1966','1967','1968','1969','1970','1971','1972','1973','1974','1975','1976','1977','1978','1979','1980','1981','1982','1983','1984','1985','1986','1987','1988','1989','1990','1991','1992','1993','1994','1995','1996','1997','1998','1999','2000','2001','2002','2003','2004',key="year",value="intake")
```

##Team

Question
Our overall question is "Has sugar intake in grams per person per day increased world wide from since 1961 until 2004?" What we are really asking is, is there a difference in the median amount of sugar consumed per person per day in 1961 versus in 2004? This is an increasingly important question, as many nations such as the United States and the United Kingdom are now facing obesity epidemics. With the Western way of life spreading around the world, this may also soon be affecting countries who are just beginning to see signs of a growing obesity problem, or may not have one yet but could reasonably see problems in the future. Answer this question will help us better humanity in the fight against obesity and health problems. If sugar intake has increased, then focusing on reducing sugar intake could lead to better overall health in the future. If there does not seem evidence that sugar intake has increased, then we are free to explore other variables that may be more useful to fighting obesity in the future. 

Null Hypothesis
We expect there will not be a difference in sugar intake for countries between the years 1961 and 2004. Therefore, we can predict the average difference in sugar intake will be around 0.

Test Statistic
The test statistic we will be using to test our hypothesis will be the median difference in sugar intake for countries between the years 1961 and 2004

Distribution of the Data
```{r}
sugardata3$year<-as.numeric(sugardata3$year)
Sugar1961<-as.tibble(filter(sugardata3, year==1961))
Sugar2004<-as.tibble(filter(sugardata3, year==2004))
full4<-full_join(Sugar1961,Sugar2004, by=c("Country"),suffix = c(".x", ".y"))
full5<-full4%>%
  mutate(diff=intake.y-intake.x)
full6<-na.omit(full5)

test<-sample(full6$diff,10)
test.median<-median(test)

null.median<-vector("double",length=1000)
for(i in seq_along(null.median)){
  temp.draw<-sample(full6$diff,10)
  null.median[i]<-median(temp.draw)
}

ggplot()+geom_histogram(aes(x=null.median))
statMedian2<-sum(null.median)/1000
```

From the distribution above, it does appear as though sugar intake has increased around the world between the years 1961 and 2004. The median is around 20, indicating an increase in sugar intake of 20 grams more per person per day in 2004 than in 1961. Thus, we can predict the real median difference in sugar intake between 1961 and 2004 has increased by about 20 grams. 

Comparing to the real statistic
```{r}
real.median2<-median(full6$diff)
percentile.data2<-sum(null.median<real.median2)/1000 #0.464
ggplot()+geom_histogram(aes(x=null.median))+geom_vline(xintercept=real.median2, color="green")
```

From the plot above we can see that the real test statistic falls around a little less than 20 grams as the median difference in sugar intake between 1961 and 2004. This is right where we would expect it to be to support our hypothesis that there has been an increase in sugar intake. The real test statistic falls at the 50th percentile, meaning the samples taken do represent the actual data. Thus, we can conclude that sugar intake did actually increase around the world between 1961 and 2004. This means decreasing the sugar consumption across the world could lead to an overall increase in health. Increasing the overall health of the world will better humanity since it could lead to a longer life expectancy and decrease instances of diseases related to obesity and weight. 

##Abby
Subquestion
We want to know if sugar intake increased for every country from 1980 to 1989. If it didn't, on average, then we would expect the difference between 1980 and 1989 to be around 0. What is the distribution of differences in sugar intake for countries between the years 1980 and 1989.

Null Hypothesis
We expect there will not be a difference in sugar intake for countries between the years 1980 and 1989. We expect the average difference in sugar intake to be around 0. 

Test Statistic
The test statistic for this test will be the median difference in sugar intake between the years 1980 and 1989. 

Distribution of the Data
```{r}
sugardata3$year<-as.numeric(sugardata3$year)
Sugar1980<-as.tibble(filter(sugardata3, year==1980))
Sugar1989<-as.tibble(filter(sugardata3, year==1989))
full<-full_join(Sugar1980,Sugar1989, by=c("Country"),suffix = c(".x", ".y"))
full2<-full%>%
  mutate(diff=intake.y-intake.x)
full3<-na.omit(full2)

test<-sample(full3$diff,10)
test.median<-median(test)

null.median<-vector("double",length=1000)
for(i in seq_along(null.median)){
  temp.draw<-sample(full3$diff,10)
  null.median[i]<-median(temp.draw)
}

ggplot()+geom_histogram(aes(x=null.median))
statMedian<-sum(null.median)/1000
```

From the distribution above, we can see that the difference in median difference of sugar intake between 1980 and 1989 seems normally distributed. Looking at the distribution above we would expect the real median to be 

Comparing to the Real Statistic
```{r}
real.median<-median(full3$diff)
percentile.data<-sum(null.median<real.median)/1000 #0.464
ggplot()+geom_histogram(aes(x=null.median))+geom_vline(xintercept=real.median, color="green")
```

From the plot above we can see visually that our distribution of medians seems to correctly approximate the real median difference in sugar intake between 1980 and 1989. Looking at the percentile of the real median, we see that the real median falls at the 46.4th percentile. Thus, roughly half the differences are less than 0 and roughly half the data is greater than 0. From this, we fail to reject the null hypothesis that there is no difference in sugar intake from 1980 to 1989. Thus, we can conclude that sugar intake did not increase worldwide from 1980 to 1989. 

##Derek

Did sugar intake increased globally from 2000 to 2004? If it didn't, on average, then we would expect the difference between 2000 and 2004 to be around 0.We expect there will not be a difference in sugar intake for countries between the years 2000 and 2004 as the null hypothesis. We expect the average difference in sugar intake to be around 0. The test statistic for this test will be the mean difference in sugar intake between the years 2000 and 2004 

```{r pressure, echo=FALSE}
library(readxl)
library(tidyverse)
sugar <- sugardata3
colnames(sugar)[1] = "country"
sugar <- sugar[order(sugar$country),]
sugar <- transform(sugar, year = as.numeric(year))
sugar <- na.omit(sugar)

```

```{r}
ggplot(sugar, aes(x=year, y=intake, col=country))+
  geom_line() +
  theme(legend.position="none")

```

```{r}

medians <- sugar %>% group_by(year) %>% summarize(med = median(intake))
means <- sugar %>% group_by(year) %>% summarize(mean = mean(intake))
mins <- sugar %>% group_by(year) %>% summarize(min = min(intake))
maxs <- sugar %>% group_by(year) %>% summarize(max = max(intake))
IQRs <- sugar %>% group_by(year) %>% summarize(iqr = IQR(intake))

ggplot()+
  geom_line(aes(x=medians$year, y=medians$med, col="Median")) +
  geom_line(aes(x=means$year, y=means$mean, col="Mean")) + 
  geom_line(aes(x=mins$year, y=mins$min, col="Min")) +
  geom_line(aes(x=maxs$year, y=maxs$max, col="Max")) +
  geom_line(aes(x=IQRs$year, y=IQRs$iqr, col="IQR")) 
```

```{r}
mixedup_consumption <- function(Country,years=c(2000,2004)) {
  diff <- filter(sugar,country==Country,year==years[2]) %>% select(intake) -
    filter(sugar,country==Country,year==years[1]) %>% select(intake) 
  diff <- as.numeric(diff) #This gets rid of the structure of diff and makes it just a number, which means that the next line is just a numeric vector.
  c(0,diff,-diff,0)
}


sugar.tbl <- tibble(country=unique(filter(sugar,year==1967)$country),a=0,b=0,c=0,d=0)

for (nm in seq_along(sugar.tbl$country)) {
  sugar.tbl[nm,2:5] <-  mixedup_consumption(sugar.tbl$country[nm])
}

null.median<-vector("double",length=1000)
for(i in seq_along(null.median)){
  temp.draw<-sample(sugar.tbl$b,10)
  null.median[i]<-mean(temp.draw)
}

ggplot()+geom_histogram(aes(x=null.median))
statMedian<-sum(null.median)/1000
sugar.tbl <- na.omit(sugar.tbl)
#sugar.tbl$b
real.median<-mean(sugar.tbl$b)
#real.median
percentile.data<-sum(null.median<real.median)/1000 #0.464
ggplot()+geom_histogram(aes(x=null.median))+geom_vline(xintercept=real.median, color="green")
```

The average mean is slightly above 0 meaning that global sugar consumption did increase from 2000 to 2004. This matches what we see on the graph of global mean sugar consumption.



##Individual Contributions
Abby: I first read in and tidied the data. First, I renamed the column variables to exclude letters from the years and include a label for the country column. Then, I gathered the data by year to reduce the number of columns from 45 to 3 variables. Then, I created a subquestion to help answer our overall question about whether sugar intake has increased worldwide since 1961. I wanted to see if sugar intake had increased during the 1980s, and tested to see if a distribution of sample means would indicate that there was or was not a difference in median sugar intake between 1980 and 1989. Then, I came to the conclusion using my data that sugar intake may not have changed during the 80s, as the test was not significantly difference to indicate otherwise. 

Derek: I read in and tidied the data, fixing column names where necessary. I then tested whether the mean consumption of sugar changes from the year 2000 to 2004 and found that it slightly increased. I also created a plot that shows the mean, median, min, max, and IQR for sugar consumption each year globally for our group question.