---
title: "Lab 15 WADDS"
author: "WADDS"
date: "May 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
library(tidyverse)
library(modelr)
library(data.table)
options(na.action = na.warn)
happiness <- read_csv("2017.csv")
```
##Team
The Question:
North Korean leader Kim Jong-Un is very interested in increasing the overall happiness of the citizens of North Korea. As a world leader, he is interested in increasing the visibility and strength of North Korea, and happy citizens tend to signify stable and powerful countries. We have been hired to determine which factors increase the happiness of citizens, so that North Korea can improve their ranking. 

We will be using data from the World Happiness Report to analyze which factors increase the overall happiness ranking of a country.
The dataset can be found here, at this link: 
https://www.kaggle.com/unsdsn/world-happiness 

```{r teampartmaybe}

fit.total<-lm(Happiness.Score~Economy..GDP.per.Capita.+Family+Health..Life.Expectancy.+Freedom+Generosity+Trust..Government.Corruption., data=happiness)
summary(fit.total)

```

The linear model describing happiness of a country is given by the equation:

$$ Y_{happiness}=1.74+0.78x_{GDP}+1.12x_{fam}+1.29x_{heatlh}+1.48x_{free}+0.38x_{gen}+0.83x_{trust}  $$

This means that for every unit increase in the economy variable, there will be an average increase in 0.78 on the happiness score when all else is held fixed. For every unit increase of the family variable, there will be an average of a 1.12 increase in the happiness score. For every unit increase in the health variable, the happiness score will increase by 1.29 on average. For every unit increase in freedom, there is an average of a 1.48 increase in the happiness score. For every unit increase in generosity, there is an average of a 0.38 increase in happiness score, and finally, for every unit increase in trust in the government, there is an average of a 0.83 increase in the happiness score when all else is held fixed. From our model, we see that 81.24 percent of the variation on global happiness can be explained by our linear model.

```{r, warning=FALSE, error=FALSE}
plot(fit.total$residuals, data=happiness)
```

From the residual plot above, we can see that there does not appear to be a pattern among the residuals. This indicates that our linear model to predict the overall happiness of a country is appropriate. Since we have created a linear model using the variables that are most strongly correlated with a country's overall happiness, we feel that we have dealt with any confounding variables and can thus use our linear model to accurately predict the overall happiness of a country.

Based on our conclusions, we recommend that North Korea increase their average life expectancy and GDP per capita. We also recommend North Korea invest in programs designed to make families feel more connected and supported by each other. In practice, this may translate to the elimination of the three generations of punishment rule, since that tends to divide families and create distrust. Although not as strongly correlated as other variables, freedom does tend to indicate higher happiness scores. If North Korea truly wants to improve their happiness ranking, we suggest switching from a dictatorship to a more democratice government to increase freedom and thus improve overal happiness.

Reflection:
Our goal for this semester is to learn how to use R to solve problems in a data science team setting. We hope to learn how to effectively communicate results and processes so we can analyze the world around us.

As a team, we feel we were able to work well in a group setting. We always found time to meet on Thursday nights to put all of our individual contributions together. We were always on time and worked well on the team parts together. As a team, we pulled together consistent and in-depth reports on our findings. And we feel we learned how to effectively communicate the results of our analysis each week. 

If we could travel back in time, we would tell our team to keep meeting every single week, and showing up on time. We should also keep reading the modules and coming to class prepared to contribute to the team tests. We would also tell ourselves to stop working individually on the team application excercises, and start acting more collaboratively on the in-class activities. We would also tell ourselves to start discussing the readings before taking the tests, so we could help each other work through any misunderstandings so we all have a greater understanding of the material. 

## Sarah
Freedom and Happiness

In this section, we will analyze how freedom will affect a general country's happiness. The freedom category of the dataset represents the extent to which freedom contributed to the overall happiness score of the country. 

First we will use a permuation hypothesis test to see if there was a significant difference in the freedom contribution for the ten happiest countries. This is done by taking a thousand random samples of the average of ten freedom contributions, and comparing the mean of the top ten with this distribution overall. 

``` {r sarah1}
null.freedom<-tibble(freedom=numeric())
for (i in 1:1000){
  samp<-sample(happiness$Freedom, 10)
  null.freedom[i,1]<-mean(samp)
}

ggplot(data=null.freedom)+geom_histogram(aes(x=freedom))+geom_vline(xintercept=mean(happiness$Freedom[1:10]), color="red")+labs(title="Mean Freedom Contributions Distribution")

percentile<-ecdf(null.freedom$freedom)
percentile(mean(happiness$Freedom[1:10]))

```

As one can see, there is indeed a significant difference between the freedom contribution in the top ten happiest countries and the rest of the countries. This suggests that freedom has a significant impact on a country's happiness. 

To further this analysis, we will perform linear regression on the relationship between the freedom contribution and the happiness scores for each country. 

```{r sarah2}

fit.freedom<-lm(Happiness.Score~Freedom, data=happiness)
summary(fit.freedom)
plot(Happiness.Score~Freedom, data=happiness, main="Happiness vs. Freedom")
abline(fit.freedom)

```

The linear model suggests that for every unit increase in the freedom contribution, the happiness score will increase by 4.3 on average, and that a country with no freedom contribution will have an estimated happiness score of 3.596. Both these pieces of the linear model are significant. However, this linear model only accounts for 32.5% of the variation in the happiness scores, so we will examine other factors as well. 

Goal Reflection

The goals I set for myself for six months and five years after graduation haven't really changed all that much with this semester. I still am hoping to graduate next year with a major in math and a focus in data science and statistics, along with a minor in creative writing. After that, I still hope to be exploring a variety of jobs or projects within jobs and getting a feel for what I might like. I also might consider getting a Masters after I get a little more experience in industry to get a better feel for what really makes me happy. I have an internship working with data this summer, and I'm hoping that will really help me solidify some goals as well. 


##Abby
Linear Modeling: 

The Question: Does generosity increase happiness? The World Happiness Report combines observational data from citizens of 155 countries each year to determine which factors increase the overall happiness of the people living in each country. Generosity is a measure of the amount of people in a given country who regularly give to charity. Generally, it seems generosity would positively impact happiness, as helping others increases our perceptions of self-worth and gives people the satisfication of knowing they have made a difference in the lives of other people. We would therefore expect a linear model analyzing the effect of generosity on happiness to indicate an increase in happiness with increasing generosity scores. 
```{r}
WHR2<-happiness
lm(Happiness.Score~Generosity, data=WHR2)
a1<-5.032
a2<-1.303

ggplot(data=WHR2) + 
  geom_abline(aes(intercept = a1, slope = a2), alpha = 1/4) +
  geom_point(aes(x=Generosity, y=Happiness.Score))
cor.test(WHR2$Generosity,WHR2$Happiness.Score)
```

The linear model fitting Happiness Scores on Generosity gives us the equation 
y = 1.303x + 5.032.
With a slope of 1.303, we can see that generosity seems to have a slightly positive effect on the happiness Score of a country. However, looking at the model visually, the correlation between generosity and happiness Scores does not seem to be very high. Testing for the correlation between generosity and happiness gives us a correlation of 0.155, which indicates a weak positive correlation between generosity and happiness scores. Taking a look at the residuals, we may be able to determine if there are possibly other confounding variables which may impact happiness. 

```{r}
gen.lm = lm(Happiness.Score~Generosity, data=WHR2) 
gen.res = resid(gen.lm)
plot(WHR2$Generosity, gen.res) 
```

Looking at the plot of the residuals, we can see there is no real pattern to the residuals, this indicates that a linear model is appropriate for this data. From the analysis above, we can see that the correlation between generosity and happiness are weakly correlated. From the plot above, however, we can also see that the linear model generated from the data is appropriate. This simply indicates that generosity does not seem to increase the overall happiness of a country's citizens, but it is still possible that generosity could impact other variables within the data set, such as health or trust in the government, which may have a stronger correlation with happiness.

Semester Reflection:

Before this class, my 6-month post-graduation goal was to be working for a larger tech company, such as HP or Intel. This plan has remained unchanged, but has been made more realistic through the course of this semester. Because of the skills I have gained through my Computer Science and statistics courses at CU, I have been offered a position as a Data Science intern at HP, where I will be working this summer. This internship is guaranteed for the duration of my time in higher education, and I have informed that most interns receive job offers upon graduation, making my 6-month goal easily attainable through hard work both this summer and any subsequent summers I decide to spend at HP.

My five year goals have also remained relatively stable, yet I do feel I have broadened my definition of what data analysis I would consider valuable to society. My work at HP will involve analyzing the likelihood of certain risks happening at various HP locations, which will increase the safety of those working there and prove to be extremely valuable information. 

In this course I learned how to visual data in a more effective and clear way. Often, it is hard to decide the best methods and practices to use in modeling data, and showing trends. I feel it has also been very helpful in hypothesis testing, as it adds a nice visual which clearly illustrates how the results of a hypothesis test inform our understanding of the data. If I were to give advice to myself at the beginning of the semester, I would have suggested looking for more outside resources to learn the material. There are plenty of professors I have taken in the past with notes I could have asked for to just clear some stuff up when it comes to R, since I struggled trying to learn it through the textbook. I would suggest that I keep staying on top of my work and getting it done before our team meets to pull all of our individual parts together. Being organized helped our team successfully complete our labs without taking too much time out of our schedules. If I were to tell myself to stop doing something, it would be to stop trying to figure out the individual assignments without studying the concepts further when I am lost. 

##David
```{r}
data <- happiness
```

__Distributions of Happiness Score and Trust Governmen Corruption__

```{r}
library(ggplot2)
ggplot(data, aes(Happiness.Score)) + geom_histogram(bins = 20, fill = "lightblue")
ggplot(data, aes(Trust..Government.Corruption.)) + geom_histogram(bins = 20, fill = "indianred")
```

The histograms shows that the distribution of Happiness Score of the 155 countries appears to be symmetric and approximate normal while the distribution of Trust Governmen Corruption of the 155 countries is seriously right-skewed. It means that the happiness score follows normal distribution but the Trust Governmen Corruption has large proportion of low scores which means lots of countries people do not trust Governmen Corruption.

__Correlation between  Happiness Score and Trust Governmen Corruption__

Now use Correlation test, the hypotheses are:

$H0: \rho = 0$ vs. $H1: \rho \neq 0$

```{r}
cor.test(data$Trust..Government.Corruption., data$Happiness.Score)
```

We test the correlation between the Happiness Score and Trust Governmen Corruption, the p value of the test isclose to 0 so that we should reject the H0 null hypothesis and conclude H1 that there is significant correlation between the Happiness Score and Trust Governmen Corruption at 5% level.

__Linear regression of the Happiness Score and Trust Governmen Corruption__

```{r}
ggplot(data, aes(Trust..Government.Corruption., Happiness.Score)) + geom_point() + geom_smooth(method = "lm")
m1 <- lm( Happiness.Score ~Trust..Government.Corruption., data = data)
summary(m1)
```

The linear regression model shows significant relation between happiness score and 
Trust Government Corruption, we can find that there is clearly ascend trend in the plot that with increasing of the Trust Government Corruption there would also increase in the happiness score. The summary output shows 1 addtional score in Trust Government Corruption is associated with 4.7746 increase in happiness score.

If I could travel back, I would discuss more with my teammates and finish my work on time. Six month after graduation, I still plan to attend graduate school and I will be majoring statistics, during this time, I should have finished 1st semester graduate study. I also have some understanding about statistics (at least for some programming skills in R). After five years, I should finish my graduate study and work in the company. In the future, I plan to do some work relate to statistics and apply my knowledge into real life,
help people solve problems. Actually these goals does not change, at least in couple of years. 

From this course I have learnt so many techniques to deal with data analysis problem by R. Compared with a lot of languages I have learnt here. I feel that R is one of the most easiest language, and it's actually a very powerful language used in Data Analysis. Although it is difficult for us to choose to best method to solve a single data science problem. It's very important for me to learn deeply so we could train ourself to solve more complex problems. If time traveled to the beginning of the semester. I would tell myself I could communicate with our teammates often, and read more books in order to prepare me for the class well. I would keep contributing my ideas and help my teammates to finish labs and projects. If I were tell myself to stop doing something, it would be like following: just stay here and do my work and submit! Be an activte person!

## Derek

```{r, echo = TRUE }
library(corrplot)
happiness <- read_csv("2017.csv")
other <- read_csv("~/Downloads/2016.csv") # I grab the older 2016 data because it also includes the region for each country
other <- other[,1:2]
drops <- c('Happiness.Rank', 'Whisker.high', 'Whisker.low') # Drop uneeded columns 
happiness <- happiness[ , !(names(happiness) %in% drops)] 
happiness <- left_join(happiness, other, by="Country") 
happiness$Country<-as.factor(happiness$Country)
# Filling in some missing data
happiness[happiness$Country == "Taiwan Province of China" ,10] <- "Eastern Asia"
happiness[happiness$Country == "Hong Kong S.A.R., China" ,10] <- "Eastern Asia"
happiness[happiness$Country == "Central African Republic" ,10] <- "Sub-Saharan Africa"
happiness[happiness$Country == "Mozambique" ,10] <- "Sub-Saharan Africa"
happiness[happiness$Country == "Lesotho" ,10] <- "Sub-Saharan Africa"
happiness$Region<-as.factor(happiness$Region)
```

I looked at whether the linear model is consistent across all types of countries. To seperate the countries into "types" I clustered them with kmeans based on the values *Family*, *Health..Life.Expectancy.*, *Freedom*, *Economy..GDP.per.Capita.*, *Generosity*, and *Trust..Government.Corruption.* into three clusters. So countries that are similar to each other will be catogrized the same. We end up with three groups that can be loosely thought of as 1st world countries, 3rd world countries and the countries inbetween. I plot these clusters with a sum of the statistics on the x-axis and Happiness.Score on the the y-axis and can see a nice positive linear relationship. We can also see that the clustering worked because we would expect countries that have similar scores to have similar happiness.

```{r}
#happiness$Health..Life.Expectancy. <- (happiness$Health..Life.Expectancy.-min(happiness$Health..Life.Expectancy.))/(max(happiness$Health..Life.Expectancy.)-min(happiness$Health..Life.Expectancy.))

#happiness$Freedom <- (happiness$Freedom-min(happiness$Freedom))/(max(happiness$Freedom)-min(happiness$Freedom))

#happiness$Family <- (happiness$Family-min(happiness$Family))/(max(happiness$Family)-min(happiness$Family))

#happiness$Economy..GDP.per.Capita. <- (happiness$Economy..GDP.per.Capita.-min(happiness$Economy..GDP.per.Capita.))/(max(happiness$Economy..GDP.per.Capita.)-min(happiness$Economy..GDP.per.Capita.))

#happiness$Trust..Government.Corruption. <- (happiness$Trust..Government.Corruption.-min(happiness$Trust..Government.Corruption.))/(max(happiness$Trust..Government.Corruption.)-min(happiness$Trust..Government.Corruption.))

#happiness$Generosity <- (happiness$Generosity-min(happiness$Generosity))/(max(happiness$Generosity)-min(happiness$Generosity))

happiness$Sum <- happiness$Freedom + happiness$Family + happiness$Health..Life.Expectancy. + happiness$Economy..GDP.per.Capita. + happiness$Trust..Government.Corruption. + happiness$Generosity
```

```{r}

k2 <- kmeans(happiness[,3:8], centers = 3, nstart = 25)
#table(k2$cluster, happiness$Region)

happiness$cluster <- k2$cluster

by_cluster <- happiness %>%
  group_by(cluster) %>%
  nest()

calc_model <- function(df) {
  lm(Happiness.Score~Sum, data=df)
}

by_cluster <- mutate(by_cluster,sum.fit=map(data,calc_model))
by_cluster
colors = as.factor(k2$cluster)

ggplot(happiness, aes(x=Sum, y=Happiness.Score, col= colors, label=Country)) +
  geom_point() +
  geom_text(aes(label=ifelse(((Happiness.Score / Sum)<1.25) & (Happiness.Score<5.5) | Country=="Syria"| Happiness.Score < 3 ,as.character(Country),'')),hjust=0,vjust=0) +  
  geom_abline(aes(intercept=by_cluster[[1,3]]$coef[[1]], slope= by_cluster[[1,3]]$coef[[2]], col = levels(colors)[1])) +
  geom_abline(aes(intercept=by_cluster[[2,3]]$coef[[1]], slope= by_cluster[[2,3]]$coef[[2]], col = levels(colors)[2])) +
  geom_abline(aes(intercept=by_cluster[[3,3]]$coef[[1]], slope= by_cluster[[3,3]]$coef[[2]], col = levels(colors)[3])) 
```

I also plotted the linear model for each cluster. Interestingly two clusters have almost identical slopes, but the cluster with the lowest scores has a noticibly smaller slope. It appears that in these countries. A question worth asking is if these countries have a lower slope because the relationship between the six statistics and hapiness changes at these lower levels or if all the countries in this 3rd group have a confounding factor. 

```{r}
res = resid(by_cluster[[1,3]])
plot(by_cluster[[1,2]]$Sum, res) 

res = resid(by_cluster[[2,3]])
plot(by_cluster[[2,2]]$Sum, res)

res = resid(by_cluster[[3,3]])
plot(by_cluster[[3,2]]$Sum, res)
```

Looking at the residuals though, there does not appear to be any patterns for any of the clusters, so a confounding variable does not seem likely and our linear model is appropriate. North Korea is not in the dataset but I highly expect the country to fall into this 3rd cluster of countries. In this case it will have have a very difficult time increasing happiness, because the increase in happiness from improving conditions is not as significant as it is for countries that are already in a better situation. 

Reflection: I havent changed my goals that I had from the begining of the semester. However over the course of the semester I have learned R and some useful data processing techniques I look forward to applying these techniques in future data related classes. If I could give myself advice at the beginning of the semester I would tell myself to focus more on time management.

##Will
GDP per Capita

In this section, we will discussion whether GDP per Capita has a significant influence in Happiness Score.

This question is interesting and important. If there is a strong positive and significant association between happiness and GDP per Capita, each government's budgeting, focus and priority will and should shift over all the world.

At the early stage, when we try to study this kind of question, the most efficient way is to load the dataset and then draw one or more plots, sometimes this method will inspire us. But only if we combine the heuristic approach like visualization with some formal tool (such as *hypothesis test* and *linear regression*) can we discover the hidden pattern. The new learned toolkit is powerful and elegant.  

Hypothesis Test Approach
First, we try to figure out whether happiness score has a significant change when GDP per capita moves from 1 unit below to the above. 
```{r}
#Load the dataset
WorldHappiness2017<-happiness
#WorldHappiness2017 <- as.tibble(read.csv( "2017.csv", sep = ",",header = TRUE, #quote = "\"",fill = TRUE))
#head(WorldHappiness2017)
#summary(WorldHappiness2017)
#colnames(WorldHappiness2017)

#Grouping by GDP per Capita, study the corresponding happiness score.
scoreHappiness_below <- filter(WorldHappiness2017, Economy..GDP.per.Capita. < 1) %>% select(Happiness.Score) %>% unlist
scoreHappiness_above <- filter(WorldHappiness2017, Economy..GDP.per.Capita. > 1) %>% select(Happiness.Score) %>% unlist

# Find the difference of means of the two groups.
obsDiff <- mean(scoreHappiness_above) - mean(scoreHappiness_below)
obsDiff
```
We assume that no matter GDP per Capita is above or below, the happiness scores are almost the same. It is our null hypothesis. To test the hypothesis, we mix up the GDP per capita labels. Then we will compare the stimulated statistics with the observed one.

```{r, echo = TRUE}
# Mix up the GDP labels
N = 75
diffAverage <- replicate(1000, {
    all <- sample(c(scoreHappiness_above,scoreHappiness_below))
    testGroup1 <- all[1:N]
    testGroup2 <- all[(N+1):(2*N)]
  return(mean(testGroup2) - mean(testGroup1))
})
hist(diffAverage)
abline(v=obsDiff, col="red", lwd=2)
```

```{r, echo = TRUE}
# Find the percentile
percentile <- ecdf(diffAverage)
percentile(obsDiff)
```
From the plot, we can see than the observed difference between the two groups lies far right from the "normal" ones. And the percentile is almost 100%. That is to say, no matter what the confidence level is, we will refuse the null hypothesis. Put in another word, happiness score will change significantly when the GDP per capita moves from 1 unit below to the above (under the confidence level 95% or 99%)

Linear Regression Approach

Now we study the happiness-GDP association problem by using linear regression
```{r, echo = TRUE}
happyGDP_model <- lm(Happiness.Score ~ Economy..GDP.per.Capita., data = WorldHappiness2017)
summary(happyGDP_model)
plot(Happiness.Score ~ Economy..GDP.per.Capita., data = WorldHappiness2017, main="Happniess and GDP per Capita")

## find the regression line
interceptHappy <- coef(summary(happyGDP_model))["(Intercept)",1]
slopeHappy <- coef(summary(happyGDP_model))["Economy..GDP.per.Capita.",1]
abline(a = interceptHappy, b = slopeHappy, col="red")
```
From the summary, we know that $R^2 = 0.6617$, which means that the variability of the happiness score can a mostly be explained by GDP per Capita. 

We can conclude that happiness of different country and GDP per Capita has as a significantly positive association. Here we use two different formal tools, we have a strong confidance about the conlusion.

Reflection on Lab 2
William Blake once wrote:

> "To see a World in a Grain of Sand 
And a Heaven in a Wild Flower, 
Hold Infinity in the palm of your hand 
And Eternity in an hour. "

I have not changed my wish to apply machine learning/deep learning tools to real world such as *finance*. 

Technique such as programming and statistical thinking have no meaning if we don't have any human understanding. The various data analyzed in computer science are standard or would put in a standard way. At the same time, it means that the data is too "thin" to make a difference. But our intuitions are always mis-guiding. We need the tools just as the ones we pick up from this course the filter the confused and wrong intuitions.

This is the first big lesion I learned from the course. Secondly, in the course and other studies, I find that more supplicated techniques and tricks are already hide in the relatively easier and thighward tools learned in this lesson. In the reflection, I am very thankful to take in the course. And I want to explore more and go far away from the origin by combining the "standardization" computer thinking and our sense-making of "thick" data.

## Summary of Individual Contributions

Sarah: I used permutation hypothesis testing to analyze whether there was a significant difference in the mean of the ten happiest countries' freedom contribution as compared with the rest of the data. I also used linear regression to determine the relationship between freedom and happiness. Finally, I reflected on the goals I set for myself during the beginning of the semester. 

Abby: I analyzed generosity and happiness to see if increased generosity lead to an increase in overall happiness for citizens of different countries. By creating a linear model, I saw there was a slightly positive correlation between generosity and happiness, but the linear model did not fit the data very well. I then created a residual plot to see if it was possible that a confounding variable would have made a linear model inappropriate. Through my analysis of the residual plot I concluded that while the linear model did not fit the data well, it was still an appropriate model. For out team part, I found the dataset we wanted to use, described our overall question and why it was important, and contributed to the analysis of our linear model. I also ploted the residuals to make sure our linear model was not being impacted by confounding variables. Finally, I reflected on this last semester and how this class has impacted me.

David: In this lab I have analyzed hapiness score and trust government corruption in the given data using linear regression, and I have used linear regression model significant relation between happiness score and Trust Government Corruption. I have also made historgram to shows that the distribution of Happiness Score of the 155 countries appears to be symmetric and approximate normal. Also I reflected what I have done for this semester and how this class affect my future career

Derek: I clustered the the data into 3 subsets based on similarity and looked at these subsets to see if the linear model is consistent accross all kinds of countries. I found that most contries share a similar linear model between the 6 statistics and happiness but countries that score very low on the 6 statistics (and therefore low hapiness) follow a different model. 

Will: For lab15, I try to discuss is whether GDP per Capital has a significant influence in Happiness Score. In order to analyze this question, I used the Null Hypothesis to figure out, whether it has statistics significant to show relationship between these two variables. Finally, happiness of different country and GDP per Capita has as a significantly positive association.