---
title: "Lab 15 Sarah"
author: "Sarah Liddle"
date: "May 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
library(tidyverse)
library(modelr)
happiness <- read_csv("happiness.csv")
```

## Freedom and Happiness

In this section, we will analyze how freedom will affect a general country's happiness. The freedom category of the dataset represents the extent to which freedom contributed to the overall happiness score of the country. 

First we will use a permuation hypothesis test to see if there was a significant difference in the freedom contribution for the ten happiest countries. This is done by taking a thousand random samples of the average of ten freedom contributions, and comparing the mean of the top ten with this distribution overall. 

``` {r sarah1}
null.freedom<-tibble(freedom=numeric())
for (i in 1:1000){
  samp<-sample(happiness$Freedom, 10)
  null.freedom[i,1]<-mean(samp)
}

ggplot(data=null.freedom)+geom_histogram(aes(x=freedom))+geom_vline(xintercept=mean(happiness$Freedom[1:10]), color="red")+labs(title="Mean Freedom Contributions Distribution")

percentile<-ecdf(null.freedom$freedom)
percentile(mean(happiness$Freedom[1:10]))

```

As one can see, there is indeed a significant difference between the freedom contribution in the top ten happiest countries and the rest of the countries. This suggests that freedom has a significant impact on a country's happiness. 

To further this analysis, we will perform linear regression on the relationship between the freedom contribution and the happiness scores for each country. 

```{r sarah2}

fit.freedom<-lm(Happiness.Score~Freedom, data=happiness)
summary(fit.freedom)
plot(Happiness.Score~Freedom, data=happiness, main="Happiness vs. Freedom")
abline(fit.freedom)

```

The linear model suggests that for every unit increase in the freedom contribution, the happiness score will increase by 4.3 on average, and that a country with no freedom contribution will have an estimated happiness score of 3.596. Both these pieces of the linear model are significant. However, this linear model only accounts for 32.5% of the variation in the happiness scores, so we will examine other factors as well. 

## Goal Reflection

The goals I set for myself for six months and five years after graduation haven't really changed all that much with this semester. I still am hoping to graduate next year with a major in math and a focus in data science and statistics, along with a minor in creative writing. After that, I still hope to be exploring a variety of jobs or projects within jobs and getting a feel for what I might like. I also might consider getting a Masters after I get a little more experience in industry to get a better feel for what really makes me happy. I have an internship working with data this summer, and I'm hoping that will really help me solidify some goals as well. 


## Summary of Individual Contributions

Sarah: I used permutation hypothesis testing to analyze whether there was a significant difference in the mean of the ten happiest countries' freedom contribution as compared with the rest of the data. I also used linear regression to determine the relationship between freedom and happiness. Finally, I reflected on the goals I set for myself during the beginning of the semester. 





