---
title: "Lab 15 Sarah"
author: "Sarah Liddle"
date: "May 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
library(tidyverse)
library(modelr)
happiness <- read_csv("happiness.csv")
```

## Freedom and Happiness

In this section, we will analyze how freedom will affect a general country's happiness. The freedom category of the dataset represents the extent to which freedom contributed to the overall happiness score of the country. 

First we will use a permuation hypothesis test to see if there was a significant difference in the freedom contribution for the ten happiest countries. This is done by taking a thousand random samples of the average of ten freedom contributions, and comparing the mean of the top ten with this distribution overall. 

``` {r sarah1}
null.freedom<-tibble(freedom=numeric())
for (i in 1:1000){
  samp<-sample(happiness$Freedom, 10)
  null.freedom[i,1]<-mean(samp)
}

ggplot(data=null.freedom)+geom_histogram(aes(x=freedom))+geom_vline(xintercept=mean(happiness$Freedom[1:10]), color="red")+labs(title="Mean Freedom Contributions Distribution")

percentile<-ecdf(null.freedom$freedom)
percentile(mean(happiness$Freedom[1:10]))

```

As one can see, there is indeed a significant difference between the freedom contribution in the top ten happiest countries and the rest of the countries. This suggests that freedom has a significant impact on a country's happiness. 

To further this analysis, we will perform linear regression on the relationship between the freedom contribution and the happiness scores for each country. 

```{r sarah2}

fit.freedom<-lm(Happiness.Score~Freedom, data=happiness)
summary(fit.freedom)
plot(Happiness.Score~Freedom, data=happiness, main="Happiness vs. Freedom")
abline(fit.freedom)

```

The linear model suggests that for every unit increase in the freedom contribution, the happiness score will increase by 4.3 on average, and that a country with no freedom contribution will have an estimated happiness score of 3.596. Both these pieces of the linear model are significant. However, this linear model only accounts for 32.5% of the variation in the happiness scores, so we will examine other factors as well. 

## Goal Reflection

The goals I set for myself for six months and five years after graduation haven't really changed all that much with this semester. I still am hoping to graduate next year with a major in math and a focus in data science and statistics, along with a minor in creative writing. After that, I still hope to be exploring a variety of jobs or projects within jobs and getting a feel for what I might like. I also might consider getting a Masters after I get a little more experience in industry to get a better feel for what really makes me happy. I have an internship working with data this summer, and I'm hoping that will really help me solidify some goals as well. 


## Summary of Individual Contributions

Sarah: I used permutation hypothesis testing to analyze whether there was a significant difference in the mean of the ten happiest countries' freedom contribution as compared with the rest of the data. I also used linear regression to determine the relationship between freedom and happiness. Finally, I reflected on the goals I set for myself during the beginning of the semester. 


```{r teampartmaybe}

fit.total<-lm(Happiness.Score~Economy..GDP.per.Capita.+Family+Health..Life.Expectancy.+Freedom+Generosity+Trust..Government.Corruption., data=happiness)
summary(fit.total)

```

The linear model describing happiness of a country is given by the equation:

$$ Y_{happiness}=1.74+0.78x_{GDP}+1.12x_{fam}+1.29x_{heatlh}+1.48x_{free}+0.38x_{gen}+0.83x_{trust}  $$

This means that for every unit increase in the economy variable, there will be an average increase in 0.78 on the happiness score when all else is held fixed. For every unit increase of the family variable, there will be an average of a 1.12 increase in the happiness score. For every unit increase in the health variable, the happiness score will increase by 1.29 on average. For every unit increase in freedom, there is an average of a 1.48 increase in the happiness score. For every unit increase in generosity, there is an average of a 0.38 increase in happiness score, and finally, for every unit increase in trust in the government, there is an average of a 0.83 increase in the happiness score when all else is held fixed. 

