---
title: "Lab 11"
author: "WADDS"
date: "April 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
library(tidyverse)
library(stringr)
library(babynames)
```

## First Three Letters of Names

The ratio of times a name begins with "Sar" in the whole dataset over the times a name contains "sar" is 0.9366. A plot below is shown of that same ratio for each given year.

Overall, the ratio of names in the dataset that start with the first three letters of Abby divided by total names that contain (in order) the first three letters of the Abby is 0.96522, or about 96.52%, which means most names with abb in them also begin with Abb. 

```{r part1sarah}
sarah1<-babynames %>% filter(str_detect(name, "^Sar"))
sarah2<-babynames %>% filter(str_detect(name, "[Ss]ar"))
#sum(sarah1$n)/sum(sarah2$n)
sarahratio1<-sarah1 %>% group_by(year) %>% summarize(startot=sum(n))
sarahratio2<-sarah2 %>% group_by(year) %>% summarize(intot=sum(n))
ratios<-inner_join(sarahratio1,sarahratio2)
ratios<-ratios %>% mutate(ratio=startot/intot)
#ggplot(data=ratios)+geom_point(aes(x=year,y=ratio))
```

```{r help}
babynames$group <- (str_extract(babynames$name, "[A-Z][a-z][a-z]"))

group_names <- babynames %>%
  subset(group %in% c("Abb","Dav","Der","Sar","Wil")) %>%
  group_by(group = str_extract(name, "[A-Z][a-z][a-z]"),year) %>%
  summarise(n = sum(n))

glimpse(group_names)
ggplot(group_names, aes(year, n)) +
  geom_line(aes(color=group)) +
  theme_bw() +
  scale_y_log10()
  ggtitle("Group names")

```


From the plot above, we can see that there is a slight decrease in names beginning with both 'Dav' and 'Wil'. However, the popularity of these names is still relatively high, which makes sense given there are lots of Wills, Williams, Davids, Daves and such in the world. So even if there is a slight decrease in the popularity of names beginning with those three letters, there are still plenty of popular names. For names beginning with 'Der', 'Sar', and 'Abb', there has been a big increase in the proportion of names beginning with those letters over the years. 'Abb' and 'Sar' names have both increased moderately, while 'Der' names have drastically increased over the years. This makes sense, as Derek is a relatively modern name, while Abbys and Sarahs and other similar names can be traced back hundreds of years, thus making them more classic. 

## Rachel and Ariel Regexs

```{r ArielRachel}
Ariel1<-filter(babynames, str_detect(babynames$name, "Ar[iy]+.l+[^a]?$"),sex=="F")
A73<-filter(Ariel1, year==1973, sex=="F")
##2 versions of Ariel in 1973
A88<-filter(Ariel1, year==1988,sex=="F")
##7 versions of Ariel in 1988
A90<-filter(Ariel1, year==1990, sex=="F")
##8 versions of Ariel in 1990

Rachel1<-filter(babynames, str_detect(babynames$name, "Ra[ey]?ch.*l"),sex=="F")
R73<-filter(Rachel1,year==1973)
##13 versions of Rachel
R88<-filter(Rachel1, year==1988)
##19 versions of Rachel
R90<-filter(Rachel1,year==1990)
##17 versions of Rachel

n73<-filter(babynames,year==1973)
chance73<-(sum(R73$n)+sum(A73$n))/sum(n73$n)
## Chances of being named Rachel or Ariel in 1973 0.00278237

n88<-filter(babynames,year==1988)
chance88<-(sum(R88$n)+sum(A88$n))/sum(n88$n)
## Chances of being named Rachel or Ariel in 1988 0.00567

n90<-filter(babynames, year==1990)
chance90<-(sum(R90$n)+sum(A90$n))/sum(n90$n)
## Chances of being named Rachel or Ariel in 1990 0.00612
```

Using the RegEx "Ar[iy]+.l+[^a]?$" we found that in 1973 there were 2 different spelling sof the name Ariel, in 1988 there were 7 different spellings, and in 1990 there were 8 different spellings of the name Ariel. This could indicate an overall increase in the popularity of the name Ariel over the years. Using the RegEx "Ra[ey]?ch.*l" we found that in 1973 there wer 13 different spellings of the name Rachel, int 1988 there were 19 different spellings, and in 1990 there were 17 different spellings of the name Rachel. This would indicate a slight increase in Rachel's around 1988, with the popularity decreasing again around 1990.

We also calculated the probability of being named either Rachel or Ariel in 1973 as around .28%, which continued to increase to .57% by 1988 and then again to .61% by 1990. This supports an overall trend observed in the increasing popularity of the name Ariel, and refelcts the slight increase in the popularity of Rachel around 1988. The likelihood of a girl being named Rachel or Ariel has increased over the years likely because of an increase in the popularity of those names.

## The Little Mermaid Effect

In this problem, we aim to find out if the movie, The Little Mermaid, preimiring in 1989 had an affect on little girls being names Ariel in the United States. In order to do this, we compared the change in proportion of girls names Ariel with the change in proportion of all girl names beginning with a vowel. 

The plot below is a histogram of the change in proportion of all girls names starting with a vowel from 1988 to 1990. The vertical line represents the change in proportion of the name Ariel specifically given the time period. 

```{r part3sarah1}
names1973<-filter(babynames, year==1973,sex=="F")
names1988<-filter(babynames, year==1988,sex=="F")
names1990<-filter(babynames, year==1990,sex=="F")

arielnames1973<-names1973 %>% filter(str_detect(name, "Ar[iy]+.l+[^a]?$"))
#sum(arielnames1973$n)/sum(names1973$n)
arielnames1988<-names1988 %>% filter(str_detect(name, "Ar[iy]+.l+[^a]?$"))
#sum(arielnames1988$n)/sum(names1988$n)
arielnames1990<-names1990 %>% filter(str_detect(name, "Ar[iy]+.l+[^a]?$"))
#sum(arielnames1990$n)/sum(names1990$n)

#(sum(arielnames1990$n)/sum(names1990$n))-(sum(arielnames1988$n)/sum(names1988$n))

vowelnames1988<-names1988 %>% filter(str_detect(name, "^[AEIOU]"))
#sum(vowelnames1988$n)/sum(names1988$n)
vowelnames1990<-names1990 %>% filter(str_detect(name, "^[AEIOU]"))
#sum(vowelnames1990$n)/sum(names1990$n)

countvowelnames1988<-vowelnames1988 %>% group_by(year,name) %>% summarize(ntot=sum(n))
countvowelnames1990<-vowelnames1990 %>% group_by(year,name) %>% summarize(ntot2=sum(n))


bothnames<-inner_join(countvowelnames1988,countvowelnames1990, by="name")
bothnames<-mutate(bothnames, prop=(ntot2/sum(names1990$n))-(ntot/sum(names1988$n)))
ggplot(data=bothnames)+geom_histogram(aes(x=prop),binwidth=.00005)+geom_vline(xintercept=0.00192)
```


The change in proportion of Ariel names was about 0.145%. As one cane see, the change in the proportion of the name Ariel between these two years is significantly higher than the majority of girl names that start with a vowel; it is definitely an outlier in all the data. This would suggest that there was something causing this significant increase in baby's being named Ariel. Since The Little Mermaid was propbably the biggest influence in pop culture of the name Ariel, it is likely that this is the cause of it. From the data, it is fair to suggest that The Little Mermaid did indeed cause an increase in babies names Ariel. 


## Our Names


A histogram of proportions of changes of all names between 2015 and 1997 and 1979 and 1997 are shown below. The red sections indicate boundaries of the histogram that we'd consider to be unusual. 

### 1997 to 2015

```{r part4sarah1}
names2015<-babynames %>% filter(year==2015)
names1997<-babynames %>% filter(year==1997)
names1979<-babynames %>% filter(year==1979)
sarah2015<-names2015 %>% filter(str_detect(name, "^Sarah?"))
sarah1997<-names1997 %>% filter(str_detect(name, "^Sarah?"))
sarah1979<-names1979 %>% filter(str_detect(name, "^Sarah?"))
#(sum(sarah2015$n)/sum(names2015$n))-(sum(sarah1997$n)/sum(names1997$n))
#(sum(sarah1997$n)/sum(names1997$n))-(sum(sarah1979$n)/sum(names1979$n))



countnames2015<-names2015 %>% group_by(year,name) %>% summarize(ntot2=sum(n))
countnames1997<-names1997 %>% group_by(year,name) %>% summarize(ntot=sum(n))
countnames1979<-names1979 %>% group_by(year,name) %>% summarize(ntot3=sum(n))

bothnames2<-inner_join(countnames2015,countnames1997, by="name")
bothnames2<-mutate(bothnames2, prop=(ntot2/sum(names2015$n))-(ntot/sum(names1997$n)))
#ggplot(data=bothnames)+geom_histogram(aes(x=prop),binwidth=0.00005)+geom_vline(xintercept=-0.00055, color="red")
bothnames3<-inner_join(countnames1979,countnames1997,by="name")
bothnames3<-mutate(bothnames3, prop=(ntot/sum(names1997$n))-(ntot3/sum(names1979$n)))


cond<-abs(bothnames2$prop)>0.0001
cond2<-abs(bothnames3$prop)>0.0001

ggplot(bothnames2,aes(x=prop))+geom_histogram(data=subset(bothnames2,cond==FALSE),fill="grey", binwidth=0.00005)+geom_histogram(data=subset(bothnames2,cond==TRUE),fill="red", binwidth=0.00005)+geom_vline(xintercept=-0.0055, color="blue")+geom_vline(xintercept=-0.00046, color="green")+labs(title="1997 to 2015")+geom_vline(xintercept=-0.00213, color="magenta")+geom_vline(xintercept=-0.000191636, color="purple")+geom_vline(xintercept=.00154, color="orange")

#percentile<-ecdf(bothnames2$prop)
#percentile(-0.0055)



```



The blue line represents the change in proportion for the name Sarah. The popularity of the name Sarah decreased by about 0.55% between the years 1997 and 2015 and decreased by about 0.14% between the years 1979 and 1997. This change in proportion represented the 0.025% percentile, so it very significantly decreased in proportion with respect to most names. 

The green line above is the change in proportion of my name between 1997 and 2015. Between 1997 and 2015 the popularity of my name has decreased by 0.046% which is below the change in proportion of most other baby names. The percentile ranking of the difference in proportion of Abby's between 1997 and 2015 is .0265, or 2.65%. This means only 2.65% of the names had a greater decrease in popularity than the name Abby for the years between 1997 and 2015.

The magenta line above is the change in proportion of the name Will between 1997 and 2015. Between 1997 and 2015 the popularity of Will has decreased by 0.21% which is below the change in proportion of most other baby names. The percentile ranking of the difference in proportion of Wills between 1997 and 2015 is .0093, or .93%. This means only .93% of the names had a greater decrease in popularity than the name Will for the years between 1997 and 2015. 

The purple line shows the change of the proportion of David's name between 1997 and 2015. During this time we have found that the name has been decreased by 0.019% which is below the change compared with other baby names. This corresponds with the 0.39% percentile.

This orange line shows the change of the proportion of Derek's name between 1979 and 1997. During this time I have found that the name has been increased by 0.154% which is above the change compared with other baby names. The number corresponds with the 99.60% percentile.

### 1979 to 1997

```{r ahhh}



ggplot(bothnames3,aes(x=prop))+geom_histogram(data=subset(bothnames3,cond2==FALSE),fill="grey", binwidth=0.00005)+geom_histogram(data=subset(bothnames3,cond2==TRUE),fill="red", binwidth=0.00005)+geom_vline(xintercept=-0.0055, color="blue")+geom_vline(xintercept=0.0003652932, color="green")+labs(title="1979 to 1997")+geom_vline(xintercept=-0.000338, color="magenta")+geom_vline(xintercept=0.000872258, color="purple")+geom_vline(xintercept=.001829341, color="orange")

```

The popularity of the name Sarah decreased by about 0.14% between the years 1979 and 1997 and is represented by the blue line in the graph. This change in proportion represented the 0.57% percentile, so it very significantly decreased in proportion with respect to most names. 
 

The green line above is the change in proportion of my name between 1979 and 1997.Between 1979 and 1997 the popularity of my name has increased by 0.037% which is above the change in proportion of most other baby names. The percentile ranking of the difference in proportion of Abby's between 1979 and 1997 is .9463, or 94.63%. This means Abby had a greater increase in popularity than 94.63% of other names for the years between 1997 and 2015.

The magenta line above is the change in proportion of the name Will between 1979 and 1997.Between 1979 and 1997 the popularity of Will has decreased by 0.03% which is below the change in proportion of most other baby names. The percentile ranking of the difference in proportion of Wills between 1979 and 1997 is .0439, or 4.39%. This means Will had a greater increase in popularity than 4.39% of other names for the years between 1997 and 2015.


This purple line shows the change of the proportion of David's name between 1979 and 1997. During this time I have found that the name has been increased by 0.087% which is above the change compared with other baby names. The number corresponds with the 0.16% percentile.

This orange line shows the change of the proportion of Derek's name between 1979 and 1997. During this time I have found that the name has been increased by 0.1829% which is above the change compared with other baby names. The number corresponds with the 99.63% percentile.

### Conclusions

In regard to gender, our team decided to be gender inclusive for the data set containing all names from 1997 to 2015 and 1979 to 1997. This way, we did not have to worry about merging different histograms, or plotting 4 different plots to analyze. With regards to the years, the majority of us (3/5) were born during 1997, so we decided to recalculate the other two based on the amount of people born with that name in 1997. This way, we did not have to try to figure out how to plot multiple histograms or how to create new data sets that included all years. From 1997 to 2015 all of our names had "significant" changes over the years. On our plot, the red parts of the histogram are where the extreme values lie. Since the difference in proportions for all of our names fall within or outside the red, all of our changes are significant. However, Sarah's name is the most significant with the largest decrease in the difference in proportion of the five of our names. From 1979 to 1997 all of our names also had "significant" changes over the years. Sarah's name, however, remained the most significant with the largest decrease in the difference in proportion of the five of our names. 

## Individual Contributions

Abby: I calculated the ratio of names beginning with the same three letters as my name to all names containing (in order) the first three letters of my name. I also calculated the change in proportion of my name between 1997 and 2015 as well as 1979 to 1997 with the percentile ranking as well. For the team part, I caculated how many different spellings of Rachel and Ariel there were for the years 1973, 1988, and 1990. I also calculated the probabilities of being named either Rachel or Ariel during those same years and analyzed what those numbers could possibly be telling us about their popularity. I also took a look at the graph Derek made for the popularity of names beginning with the same first three letters of each of our names and analyzed what the trends meant. 

Sarah: I calculated the ratio of names that started with "Sar" in relation to all names that contained "sar" in the dataset. Then, I found out how many different variations of the names Rachel and Ariel were in each given years. I also plotted the change in proprtion of the name Ariel before and after The Little Mermaid came out in relation to the change in proportions for all babynames in this time. Finally, I created plots of changes in proportions for all baby names from 1997 to 2015 and 1979 to 1997 and put lines for all my team members proportions of change on the plot. 

David: Here, I have implemented the factors and strings to answer some of questions in the lab. First, I have calculated the ratio of my name from the whole babyname data set, which is 0.934. After that, I have used the knowledge from factors and strings to calculate the number of names of Ariel and Rachel appear in the dataset. Then, I implemented the ratio from 1988 to 1990 to calculate the change of Ariel is -0.00177116. Therefore, we could conclude that the Little Mermaid cause less baby girls to be named Ariel. I have drawn a histogram about the Ariel among all the female names and found that Ariel has been placed in the 100 percentile change so it means that it has biggest change. Finally, I have implemented my name "David" to calculate the name change from 1979 to 1997 and 1997 to 2015. I have drawn a histogram to show the change of my name in these years. My name has been decreased by 0.019% from 1997 and 2015. My name has been increased by 0.087% between 1979 and 1997 

Will:For the first part, I used a lot str_detect and length to determine the different version of the names. In 1988 and 1900, we find 6 different versions of the "Ariel" and 4 differnet versions for "Rachel". For the second part, I take name "Tom" and "Sam" as two referene names, we can have an educated guess that  the change in the proportion of Ariels is medium compared to the changes in other female names over that time period. For the third part, I calculated a lot probability that reflects the change from 1979 to 1997 and 1997 to 2015. The result is the "will" is in the "tails". The popularity of "will" decreased consistantly.


Derek: I created the plot of the frequency of all the names that start with the first three letters of every name in our group. I did this by first creating a new group column for every name based on the first three letters of the name and then grouping by that group and the year and summing the counts of those groups. I also did some of the other misc calculations for frequency of names.