---
title: "Will_Lab 11"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(babynames)
knitr::opts_chunk$set(echo = TRUE)
```

## Ariel and Rachel regexs

```{r, echo = TRUE}
# Baby name of girl born in 1973
babynamesF1973 <- as.tibble(filter(babynames, sex == "F", year == 1973))
num_Female1973 <- sum(babynamesF1973$n, dims =2)

# Baby name of girl born in 1988
babynamesF1988 <- as.tibble(filter(babynames, sex == "F", year == 1988))
num_Female1988 <- sum(babynamesF1988$n, dims =2)

# Baby name of girl born in 1989
babynamesF1989 <- as.tibble(filter(babynames, sex == "F", year == 1989))
num_Female1989 <- sum(babynamesF1989$n, dims =2)

# Baby name of girl born in 1990
babynamesF1990 <- as.tibble(filter(babynames, sex == "F", year == 1990))
num_Female1990 <- sum(babynamesF1990$n,dim =2)

# In 1973,Female baby names with different versions of "Ariel"
pattern_ariel1973 <- str_detect(babynamesF1973$name, "^Ariel")
Ariel1973 <- babynamesF1973[pattern_ariel1973,]
unique(Ariel1973$name)
versions_Ariel1973 <- length(unique(Ariel1973$name))
num_Ariel1973 <- sum(Ariel1973$n, dims =2)
prob_Ariel1973 <- num_Ariel1973/num_Female1973

# Number of versions of female name Ariel
versions_Ariel1973

# In 1988,Female baby names with different versions of "Ariel"
pattern_ariel1988 <- str_detect(babynamesF1988$name, "^Ariel")
Ariel1988 <- babynamesF1988[pattern_ariel1988,]
unique(Ariel1988$name)
versions_Ariel1988 <- length(unique(Ariel1988$name))
num_Ariel1988 <- sum(Ariel1988$n, dims =2)
prob_Ariel1988 <- num_Ariel1988/num_Female1988

# Number of versions of female name Ariel
versions_Ariel1988


# In 1989,Female baby names with different versions of "Ariel"
pattern_ariel1989 <- str_detect(babynamesF1989$name, "^Ariel")
Ariel1989 <- babynamesF1989[pattern_ariel1989,]
# unique(Ariel1989$name)
# versions_Ariel1989 <- length(unique(Ariel1989$name))
num_Ariel1989 <- sum(Ariel1989$n, dims =2)
# prob_Ariel1989 <- num_Ariel1989/num_Female1989

# Number of versions of female name Ariel
versions_Ariel1988

# In 1990,Female baby names with different versions of "Ariel"
pattern_ariel1990 <- str_detect(babynamesF1990$name, "^Ariel")
Ariel1990 <- babynamesF1990[pattern_ariel1990,]
unique(Ariel1990$name)
versions_Ariel1990 <- length(unique(Ariel1990$name))
num_Ariel1990 <- sum(Ariel1990$n, dims =2)
prob_Ariel1990 <- num_Ariel1990/num_Female1990

# Number of versions of female name Ariel
versions_Ariel1990

## output_prob_Ariel <- c(prob_Ariel1973,prob_Ariel1988,prob_Ariel1990)
## output_prob_Ariel
```

So, there 3, 6  and 6 different versions of "Ariel" in the year 1973, 1988 and 1990 respectively.


```{r, echo = TRUE}
# In 1973,Female baby names with different versions of "Rachel"
pattern_rachel1973 <- str_detect(babynamesF1973$name, "^Rachel")
Rachel1973 <- babynamesF1973[pattern_rachel1973,]
## Rachel1973
unique(Rachel1973$name)
versions_Rachel1973 <- length(unique(Rachel1973$name))
num_Rachel1973 <- sum(Rachel1973$n, dims =2)
prob_Rachel1973 <- num_Rachel1973/num_Female1973


# Number of versions of female name Racel
versions_Rachel1973

# In 1988,Female baby names with different versions of "Rachel"
pattern_rachel1988 <- str_detect(babynamesF1988$name, "^Rachel")
Rachel1988 <- babynamesF1988[pattern_rachel1988,]
## Rachel1988
unique(Rachel1988$name)
versions_Rachel1988 <- length(unique(Rachel1988$name))
Rachel1988 <- filter(Rachel1973, !(name %in% c("Rachelanne","Rachelann","Rachelleann")))
num_Rachel1988 <- sum(Rachel1988$n, dims =2)
prob_Rachel1988 <- num_Rachel1988/num_Female1988

# Number of versions of female name Rachel
versions_Rachel1988

# In 1990,Female baby names with different versions of "Rachel"
pattern_rachel1990 <- str_detect(babynamesF1990$name, "^Rachel")
Rachel1990 <- babynamesF1990[pattern_rachel1990,]
## Rachel1990
unique(Rachel1990$name)
versions_Rachel1990 <- length(unique(Rachel1990$name))
Rachel1988 <- filter(Rachel1973, !(name %in% c("Rachelann","Rachelmarie")))
num_Rachel1990 <- sum(Rachel1988$n, dims =2)
prob_Rachel1990 <- num_Rachel1990/num_Female1990

# Number of versions of female name Rachel
versions_Rachel1990

output_prob <- c(prob_Ariel1973+prob_Rachel1973,prob_Ariel1988+prob_Rachel1988,prob_Ariel1990+prob_Rachel1990)
output_prob
```
So, there 5, 4  and 4 different versions of "Ariel" in the year 1973, 1988 and 1990 respectively. Here we find that the name ended with "n" or "ne" and "RachelMarie" have no common orign with that of "Rachel".

The chances a girl born in 1973 would be named either Rachel or Ariel (including various versions) are 0.49%, 0.50% and 0.66% in corresponding year.


## *The Little Mermaid* Effect

*The Little Mermaid* was released in the US in 1989. We want to eplore how much the release of the film affect the behavior of name-taking.

To answer the question our clients are intersted inm *i.e.* "What was the change in proportion in female Ariel baby names", we will consider changes in female names starting with a vowel from 1988 to 1990.

```{r, echo = TRUE}
# Female names starting with a vowel in various years 1988~1990
# 1988
pattern_vowel1988 <- str_detect(babynamesF1988$name, "^[AEIOU]")
vowel_babynames1988 <- babynamesF1988[pattern_vowel1988,]
num_withVowel1988 <- sum(vowel_babynames1988$n, dims=2)

# 1989
pattern_vowel1989 <- str_detect(babynamesF1989$name, "^[AEIOU]")
vowel_babynames1989 <- babynamesF1989[pattern_vowel1989,]
num_withVowel1989 <- sum(vowel_babynames1989$n, dims=2)

# 1990
pattern_vowel1990 <- str_detect(babynamesF1990$name, "^[AEIOU]")
vowel_babynames1990 <- babynamesF1990[pattern_vowel1990,]
num_withVowel1990 <- sum(vowel_babynames1990$n, dims=2)

diff_WithVowel <- num_withVowel1990 - num_withVowel1988
diff_Ariel <- num_Ariel1990 - num_Ariel1988
Ratio_Vowel <- diff_Ariel/diff_WithVowel

# In context of name "Tom"
# 1988
pattern_Tom1988 <- str_detect(babynamesF1988$name, "^Tom")
Tom1988 <- babynamesF1988[pattern_Tom1988,]
num_Tom1988 <- sum(Tom1988$n, dims=2)

# 1990
pattern_Tom1990 <- str_detect(babynamesF1990$name, "^Tom")
Tom1990 <- babynamesF1990[pattern_Tom1990,]
num_Tom1990 <- sum(Tom1990$n, dims=2)

diff_Tom <- num_Tom1990 - num_Tom1988
Ratio_Tom <- diff_Ariel/diff_Tom


# In the context of name "Sam"
pattern_Sam1988 <- str_detect(babynamesF1988$name, "^Sam")
Sam1988 <- babynamesF1988[pattern_Sam1988,]
num_Sam1988 <- sum(Sam1988$n, dims=2)

# 1990
pattern_Sam1990 <- str_detect(babynamesF1990$name, "^Sam")
Sam1990 <- babynamesF1990[pattern_Sam1990,]
num_Sam1990 <- sum(Sam1990$n, dims=2)

diff_Sam <- num_Sam1990 - num_Sam1988
Ratio_Sam <- diff_Ariel/diff_Sam
Ratio_Sam
c(diff_Sam,diff_Ariel,diff_Tom)
```
Considering all the changes in female names starting with a vowel from 1988 to 1990, Ariel names increased with about 13.1%?


```{r , echo=TRUE}
df <- data.frame(x = c("Vowel", "Tom", "Sam"), y = c(Ratio_Vowel,Ratio_Tom,Ratio_Sam))
ggplot(df, aes(x,y)) + geom_point()
```

We take name "Tom" and "Sam" as two referene names, we can have an educated guess that  the change in the proportion of Ariels is medium compared to the changes in other female names over that time period.

## Our names
```{r, echo=TRUE}
babynames1979 <- as.tibble(filter(babynames, year == 1979))
num1979 <- sum(babynames1979$n, dims = 2)

babynames1997 <- as.tibble(filter(babynames, year == 1997))
num1997 <- sum(babynames1997$n, dims = 2)

babynames2015 <- as.tibble(filter(babynames, year == 2015))
num2015 <- sum(babynames2015$n, dims = 2)

pattern1_will1979 <- str_detect(babynames1979$name,"^Wil")
will1_1979 <- babynames1979[pattern1_will1979, ]
num_will1_1979 <- sum(will1_1979$n, dims = 2)

pattern1_will1997 <- str_detect(babynames1997$name,"^Wil")
will1_1997 <- babynames1997[pattern1_will1997, ]
num_will1_1997 <- sum(will1_1997$n, dims = 2)

pattern1_will2015 <- str_detect(babynames2015$name,"^Wil")
will1_2015 <- babynames2015[pattern1_will2015, ]
num_will1_2015 <- sum(will1_2015$n, dims = 2)

pattern2_will1979 <- str_detect(babynames1979$name,"ill$")
will2_1979 <- babynames1979[pattern2_will1979, ]
num_will2_1979 <- sum(will2_1979$n, dims =2)

pattern2_will1997 <- str_detect(babynames1997$name,"ill$")
will2_1997 <- babynames1997[pattern2_will1997, ]
num_will2_1997 <- sum(will2_1997$n, dims =2)

pattern2_will2015 <- str_detect(babynames2015$name,"ill$")
will2_2015 <- babynames2015[pattern2_will2015, ]
num_will2_2015 <- sum(will2_2015$n, dims =2)

diff1979 <- num1997 -num1979
diff2015 <- num2015 - num1997

diff1_1979 <- num_will1_1997 - num_will1_1979
Ratio1_1979 <- diff1_1979/num_will1_1979
relRatio1_1979 <- diff1_1979/diff1979
diff1_2015 <- num_will1_2015 - num_will1_1997
Ratio1_2015 <- diff1_2015/num_will1_1997
relRatio1_2015 <- diff1_2015/diff2015


diff2_1979 <- num_will2_1997 - num_will2_1979
Ratio2_1979 <- diff2_1979/num_will2_1979
relRatio2_1979 <- diff2_1979/diff1979
diff2_2015 <- num_will2_2015 - num_will2_1997
Ratio2_2015 <- diff2_2015/num_will2_1997
relRatio2_2015 <- diff2_2015/diff2015
c(Ratio1_2015,Ratio1_1979,Ratio2_2015,Ratio2_1979,Ratio1_2015,Ratio1_1979,relRatio2_2015,relRatio2_1979)
```

From the statistics computed above, we can see that the popularity of "Will" decreased consistantly.

```{r, echo=TRUE}
df1 <- data.frame(x= c("1st Period","2nd Period"),y=c(relRatio1_1979,relRatio1_2015)) 
ggplot(df1,aes(x,y))+geom_col()  
df2 <- data.frame(x= c("1st Period","2nd Period"),y=c(relRatio2_1979,relRatio2_2015)) 
ggplot(df2,aes(x,y))+geom_col()  


```

From the plotting above, we find that in the first period(1979~1997) of the second version of "Will" is in the "Tails".

## Contribution
For the first part, I used a lot str_detect and length to determine the different version of the names. In 1988 and 1900, we find 6 different versions of the "Ariel" and 4 differnet versions for "Rachel". For the second part, I take name "Tom" and "Sam" as two referene names, we can have an educated guess that  the change in the proportion of Ariels is medium compared to the changes in other female names over that time period. For the third part, I calculated a lot probability that reflects the change from 1979 to 1997 and 1997 to 2015. The result is the "will" is in the "tails". The popularity of "will" decreased consistantly.
