---
title: "Will_Lab 12 0412"
output: html_document
---

```{r setup, include=FALSE}
library(gapminder) # data from gapminder.org
library(dplyr) # used to perform data transformation and manipulation
library(ggplot2) # used for data visualization
knitr::opts_chunk$set(echo = TRUE)
```

## The Trend of Life Expectancy
We find the question interesing whether the religeon affects the life expectancy. From the analysis below, we find that the life expecancy decreasing by the order of Oceania, Europe, Americas, Asia and Arican. From the plots below, we can see that the measurement such as mean is significant.
```{r, echo = TRUE}
lifeExp_Oceania <- filter(gapminder,continent=="Oceania") %>% select(lifeExp) %>% unlist
lifeExp_Europe <- filter(gapminder,continent=="Europe") %>% select(lifeExp) %>% unlist
lifeExp_Americas <- filter(gapminder,continent=="Americas") %>% select(lifeExp) %>% unlist
lifeExp_Asia  <- filter(gapminder,continent=="Asia") %>% select(lifeExp) %>% unlist
lifeExp_Africa <- filter(gapminder,continent=="Africa") %>% select(lifeExp) %>% unlist
aggregate(lifeExp ~ continent, gapminder, min)
obsdiff0_min <- 69.120 - 43.585
obsdiff1_min <- 43.585 - 37.579
obsdiff2_min <- 37.579 - 28.801
obsdiff3_min <- 28.801 - 23.599
aggregate(lifeExp ~ continent, gapminder, max)
obsdiff0_max <- 81.235 - 81.757
obsdiff1_max <- 81.757 - 80.653
obsdiff2_max <- 80.653 - 82.603
obsdiff3_max <- 82.603 - 76.442
aggregate(lifeExp ~ continent, gapminder, mean)
obsdiff0_mean <- 74.32621 - 71.90369
obsdiff1_mean <-71.90369 - 64.65874
obsdiff2_mean <- 64.65874 - 60.06490
obsdiff3_mean <- 60.06490 - 48.86533
aggregate(lifeExp ~ continent, gapminder, median)
obsdiff0_median <- 73.6650 - 72.2410
obsdiff1_median <- 72.2410 - 67.0480
obsdiff2_median <- 67.0480 - 61.7915
obsdiff3_median <- 61.7915 - 47.7920
aggregate(lifeExp ~ continent, gapminder, IQR)
obsdiff0_IQR <- 6.3475 - 5.8805
obsdiff1_IQR <- 5.8805 - 13.2895
obsdiff2_IQR <- 13.2895 - 18.0790
obsdiff3_IQR <- 18.0790 - 12.0390

N = 12
diff_lifeExp0 <- replicate(1000, {
    all <- sample(c(lifeExp_Oceania,lifeExp_Europe))
    newlifeExp_Oceania <- all[1:N]
    newlifeExp_Europe <- all[(N+1):(2*N)]
  return(summary(lifeExp_Oceania) - summary(newlifeExp_Europe))
})
hist(diff_lifeExp0["Min.",])
abline(v=obsdiff0_min, col="red", lwd=2)
hist(diff_lifeExp0["Max.",])
abline(v=obsdiff0_min, col="red", lwd=2)
hist(diff_lifeExp0["Mean",])
abline(v=obsdiff0_min, col="red", lwd=2)
hist(diff_lifeExp0["Median",])
abline(v=obsdiff0_min, col="red", lwd=2)

diff_lifeExp1 <- replicate(1000, {
    all <- sample(c(lifeExp_Europe,lifeExp_Americas))
    newlifeExp_Europe <- all[1:N]
    newlifeExp_Americas <- all[(N+1):(2*N)]
  return(summary(lifeExp_Europe) - summary(newlifeExp_Americas))
})
hist(diff_lifeExp1["Min.",])
abline(v=obsdiff1_min, col="red", lwd=2)
hist(diff_lifeExp0["Max.",])
abline(v=obsdiff1_max, col="red", lwd=2)
hist(diff_lifeExp0["Mean",])
abline(v=obsdiff1_mean, col="red", lwd=2)
hist(diff_lifeExp0["Median",])
abline(v=obsdiff1_median, col="red", lwd=2)

diff_lifeExp2 <- replicate(1000, {
    all <- sample(c(lifeExp_Americas,lifeExp_Asia))
    newlifeExp_Americas <- all[1:N]
    newlifeExp_Asia <- all[(N+1):(2*N)]
  return(summary(lifeExp_Americas) - summary(newlifeExp_Asia))
})
hist(diff_lifeExp2["Min.",])
abline(v=obsdiff2_min, col="red", lwd=2)
hist(diff_lifeExp0["Max.",])
abline(v=obsdiff2_max, col="red", lwd=2)
hist(diff_lifeExp0["Mean",])
abline(v=obsdiff2_mean, col="red", lwd=2)
hist(diff_lifeExp0["Median",])
abline(v=obsdiff2_median, col="red", lwd=2)

diff_lifeExp3 <- replicate(1000, {
    all <- sample(c(lifeExp_Asia,lifeExp_Africa))
    newlifeExp_Asia <- all[1:N]
    newlifeExp_Africa <- all[(N+1):(2*N)]
  return(summary(lifeExp_Asia) - summary(newlifeExp_Africa))
})
hist(diff_lifeExp3["Min.",])
abline(v=obsdiff1_min, col="red", lwd=2)
hist(diff_lifeExp3["Max.",])
abline(v=obsdiff1_max, col="red", lwd=2)
hist(diff_lifeExp3["Mean",])
abline(v=obsdiff1_mean, col="red", lwd=2)
hist(diff_lifeExp3["Median",])
abline(v=obsdiff1_median, col="red", lwd=2)

```


## The trend of GDP per captia
We find the question interesing whether the religeon affects the local GDP per capita. From the analysis below, we find that the quantity decreasing by the order of Oceania, Europe, Asia, Americas, and Arican. From the plots below and statistics, we come to a conclustion that the related measurement is too votile so that the result is unreliable. From the plots below,we can see that the measurement such as mean is significant but "Min" is not.

```{r, echo = TRUE}
GDP_Oceania <- filter(gapminder,continent=="Oceania") %>% select(gdpPercap ) %>% unlist
GDP_Europe <- filter(gapminder,continent=="Europe") %>% select(gdpPercap ) %>% unlist
GDP_Asia  <- filter(gapminder,continent=="Asia") %>% select(gdpPercap ) %>% unlist
GDP_Americas <- filter(gapminder,continent=="Americas") %>% select(gdpPercap ) %>% unlist
GDP_Africa <- filter(gapminder,continent=="Africa") %>% select(gdpPercap ) %>% unlist
aggregate(gdpPercap  ~ continent, gapminder, min)
gdp_obsdiff0_min <- 10039.5956 - 973.5332
gdp_obsdiff1_min <- 973.5332 - 331.0000
gdp_obsdiff2_min <- 331.0000 - 1201.6372
gdp_obsdiff3_min <- 1201.6372 - 241.1659
aggregate(gdpPercap  ~ continent, gapminder, max)
gdp_obsdiff0_max <- 34435.37 - 49357.19
gdp_obsdiff1_max <- 49357.19 - 113523.13
gdp_obsdiff2_max <- 113523.13 - 42951.65
gdp_obsdiff3_max <- 42951.65 - 21951.21
aggregate(gdpPercap  ~ continent, gapminder, mean)
gdp_obsdiff0_mean <- 18621.609 - 14469.476
gdp_obsdiff1_mean <-14469.476 - 7902.150
gdp_obsdiff2_mean <- 7902.150 - 7136.110
gdp_obsdiff3_mean <- 7136.110 - 2193.755
aggregate(gdpPercap  ~ continent, gapminder, median)
gdp_obsdiff0_median <- 8072.258 - 13248.301
gdp_obsdiff1_median <- 13248.301 - 7492.262
gdp_obsdiff2_median <- 7492.262 - 4402.431
gdp_obsdiff3_median <- 4402.431 - 1616.170
aggregate(gdpPercap  ~ continent, gapminder, IQR)
gdp_obsdiff0_IQR <- 8072.258 - 13248.301
gdp_obsdiff1_IQR <- 13248.301 - 7492.262
gdp_obsdiff2_IQR <- 7492.262 -4402.431
gdp_obsdiff3_IQR <- 4402.431 - 1616.170

N = 12
diff_GDP0 <- replicate(1000, {
    all <- sample(c(lifeExp_Oceania,lifeExp_Europe))
    newGDP_Oceania <- all[1:N]
    newGDP_Europe <- all[(N+1):(2*N)]
  return(summary(GDP_Oceania) - summary(newGDP_Europe))
})
hist(diff_GDP0["Min.",])
abline(v=gdp_obsdiff0_min, col="red", lwd=2)
hist(diff_GDP0["Max.",])
abline(v=gdp_obsdiff0_min, col="red", lwd=2)
hist(diff_GDP0["Mean",])
abline(v=gdp_obsdiff0_min, col="red", lwd=2)
hist(diff_GDP0["Median",])
abline(v=gdp_obsdiff0_min, col="red", lwd=2)

diff_GDP1 <- replicate(1000, {
    all <- sample(c(GDP_Europe,GDP_Americas))
    newGDP_Europe <- all[1:N]
    newGDP_Americas <- all[(N+1):(2*N)]
  return(summary(GDP_Europe) - summary(newGDP_Americas))
})
hist(diff_GDP1["Min.",])
abline(v=gdp_obsdiff1_min, col="red", lwd=2)
hist(diff_GDP0["Max.",])
abline(v=gdp_obsdiff1_max, col="red", lwd=2)
hist(diff_GDP0["Mean",])
abline(v=gdp_obsdiff1_mean, col="red", lwd=2)
hist(diff_GDP0["Median",])
abline(v=gdp_obsdiff1_median, col="red", lwd=2)

diff_GDP2 <- replicate(1000, {
    all <- sample(c(GDP_Americas,GDP_Asia))
    newGDP_Americas <- all[1:N]
    newGDP_Asia <- all[(N+1):(2*N)]
  return(summary(GDP_Americas) - summary(newGDP_Asia))
})
hist(diff_GDP2["Min.",])
abline(v=gdp_obsdiff2_min, col="red", lwd=2)
hist(diff_GDP0["Max.",])
abline(v=gdp_obsdiff2_max, col="red", lwd=2)
hist(diff_GDP0["Mean",])
abline(v=gdp_obsdiff2_mean, col="red", lwd=2)
hist(diff_GDP0["Median",])
abline(v=gdp_obsdiff2_median, col="red", lwd=2)

diff_GDP3 <- replicate(1000, {
    all <- sample(c(GDP_Asia,GDP_Africa))
    newGDP_Asia <- all[1:N]
    newGDP_Africa <- all[(N+1):(2*N)]
  return(summary(GDP_Asia) - summary(newGDP_Africa))
})
hist(diff_GDP3["Min.",])
abline(v=gdp_obsdiff1_min, col="red", lwd=2)
hist(diff_GDP3["Max.",])
abline(v=gdp_obsdiff1_max, col="red", lwd=2)
hist(diff_GDP3["Mean",])
abline(v=gdp_obsdiff1_mean, col="red", lwd=2)
hist(diff_GDP3["Median",])
abline(v=gdp_obsdiff1_median, col="red", lwd=2)

```





```{r pressure, echo=FALSE}
 # ggplot(gapminder_unfiltered, 
#        aes(year,gdpPercap, colour = continent)) + 
 #    geom_smooth() + ggtitle("GDP over Time by Continent") +
  #   xlab("Year") + ylab("Avg. GDP per Capita") +
   #  theme_classic()
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Contribution
For the life expeectancy, we find the question interesing whether the religeon affects the life expectancy. From the analysis below, we find that the life expecancy decreasing by the order of Oceania, Europe, Americas, Asia and Arican. From the plots below, we can see that the measurement such as mean is significant. For the gdp, we find that the quantity decreasing by the order of Oceania, Europe, Asia, Americas, and Arican. From the plots below and statistics, we come to a conclustion that the related measurement is too votile so that the result is unreliable. 